---
title: "EDA and IDF-modeling"
author: "Daniel Cheung"
date: "2026-04-09"
output: 
  pdf_document:
    number_sections: true
---

# Preperation

```{r, message = F, warning = F}
if (!requireNamespace("pacman", quietly = TRUE)) {
  install.packages("pacman")
}
library(pacman)

# Load necessary libraries
p_load(caret)
p_load(tidyverse)
p_load(tidymodels)
p_load(tidytext)
p_load(textrecipes)
p_load(Matrix)
p_load(wordcloud)
p_load(tm)

p_load(stopwords)
p_load(RColorBrewer)

p_load(glmnet)
p_load(pROC)


```



```{r}
# ---------------------------------------------------------
# 1. Data Preparation
# ---------------------------------------------------------

df <- read_csv("../../data/processed/glassdoor_reviews_cleaned_v2.csv")
# df <- read_csv("data/processed/glassdoor_reviews_cleaned_v2.csv")

cols <- c('summary', 'advice_to_management', 'review_pros', 'review_cons')
encode_cols <- c('covid_related_review', 'flags_business_outlook', 'flags_ceo_approval',
                 'flags_recommend_frend', 'rating_culture_values',
                 'rating_diversity_inclusion', 'rating_overall', 'rating_work_life',
                 'rating_compensation_benefits', 'rating_senior_leadership',
                 'rating_career_opportunities',
                 'has_review_text', 'career_missing')
df <- df %>%
  mutate(across(all_of(cols), ~replace_na(., "")))
# Define the columns

```

In the preliminary phase of the Exploratory Data Analysis (EDA), a word cloud was generated to visualize the most prominent terms within the comment dataset. To ensure the visualization reflected meaningful semantic patterns, the initial stopword list was expanded to exclude high-frequency terms that lacked domain-specific informational value.

```{r, message = F, warning = F}

# 1. Define custom stopwords
# Equivalent to STOPWORDS.union(set([...]))
custom_stopwords <- c(stopwords("english"), 
                      "bad", "good", "great", "will", "company", 
                      "job", "work", "employee", "here", 
                      "people", "employees")

# 2. Define the plotting function
plotWordCloudFromDF <- function(df, column_name) {
  
  # Extract text and collapse into one large string
  text_data <- paste(df[[column_name]], collapse = " ")
  
  # Create a Corpus and clean the text
  docs <- Corpus(VectorSource(text_data))
  docs <- docs %>%
    tm_map(content_transformer(tolower)) %>%
    tm_map(removePunctuation) %>%
    tm_map(removeNumbers) %>%
    tm_map(removeWords, custom_stopwords) %>%
    tm_map(stripWhitespace)
  
  # Generate the Word Cloud
  cat(paste0("Generating word cloud for: ", column_name, "\n"))
  wordcloud(words = docs, 
            max.words = 300, 
            random.order = FALSE, 
            rot.per = 0.35,            # Rotation percentage
            colors = brewer.pal(8, "Dark2"),
            scale = c(3, 0.5))        # Range of font sizes
}

# 3. Loop through columns


for (col in cols) {
  plotWordCloudFromDF(df, col)
}

```



The above is the data preperation step, mix up TF-IDF, dummy variable and dataset split.
```{r, message = F, warning = F}
# Split Data (equivalent to train_test_split)
set.seed(42)
df2 = df %>%
  subset(select = -c(rating_date, employee_status, company_name, 
  employee_location, employee_job_title, sector, employee_job_end_year,
  review_id))


# Build the Preprocessing Recipe (Equivalent to ColumnTransformer + TfidfVectorizer)
rec <- recipe(employee_length + employee_job_quit ~ ., data = df2)  %>%
  step_mutate(across(all_of(encode_cols), as.factor)) %>%
  # Handle Numeric/Categorical preprocessing
  step_novel(all_of(encode_cols)) %>% 
  step_dummy(all_of(encode_cols), one_hot = FALSE) %>%
  
  # Text Processing for 'summary' (max_features = 200)
  step_tokenize(summary) %>%
  step_stopwords(summary) %>%
  step_ngram(summary, min_num_tokens = 1, num_tokens = 3) %>%
  step_tokenfilter(summary, max_tokens = 200, min_times = 5) %>%
  step_tfidf(summary, prefix = "") %>%
  
  # Text Processing for 'advice_to_management' (max_features = 200)
  step_tokenize(advice_to_management) %>%
  step_stopwords(advice_to_management) %>%
  step_ngram(advice_to_management, min_num_tokens = 1, num_tokens = 3) %>%
  step_tokenfilter(advice_to_management, max_tokens = 200, min_times = 5) %>%
  step_tfidf(advice_to_management, prefix = "") %>%
  
  # Text Processing for 'review_pros' (max_features = 300)
  step_tokenize(review_pros) %>%
  step_stopwords(review_pros) %>%
  step_ngram(review_pros, min_num_tokens = 1, num_tokens = 3) %>%
  step_tokenfilter(review_pros, max_tokens = 300, min_times = 5) %>%
  step_tfidf(review_pros, prefix = "") %>%
  
  # Text Processing for 'review_cons' (max_features = 400)
  step_tokenize(review_cons) %>%
  step_stopwords(review_cons) %>%
  step_ngram(review_cons, min_num_tokens = 1, num_tokens = 3) %>%
  step_tokenfilter(review_cons, max_tokens = 400, min_times = 5) %>%
  step_tfidf(review_cons, prefix = "") %>%
  step_zv(all_predictors()) %>%
  # Scaling (Equivalent to MaxAbsScaler)
  # R's step_normalize is Z-score; for MaxAbs we use a custom scaling or range
  step_mutate(across(where(is.numeric), ~replace_na(., 0))) %>%
  step_range(all_predictors(), min =0, max = 1)


set.seed(42)
data_split <- df2$employee_job_quit%>%
  createDataPartition(p = 0.8, list = F)
train_data <- df2[data_split, ]
test_data  <- df2[-data_split, ]

# Train the recipe on training data
prep_rec <- prep(rec, training = train_data)

# Transform the datasets (Equivalent to fit_transform / transform)
X_train_processed <- bake(prep_rec, new_data = train_data, all_predictors(), composition = "dgCMatrix")
X_test_processed  <- bake(prep_rec, new_data = test_data, all_predictors(), composition = "dgCMatrix")

# Target vectors
yl_train <- train_data$employee_length
yq_train <- as.integer(train_data$employee_job_quit)

yl_test <- test_data$employee_length
yq_test <- as.integer(test_data$employee_job_quit)

# --- Data Structural Audit ---
cat("--- Data Structural Audit ---\n")
cat(sprintf("Total Combined Features: %d\n", ncol(X_train_processed)))
cat(sprintf("Total Samples: %d\n", nrow(X_train_processed)))
cat("First 30 Feature Names (Sample):\n")
print(colnames(X_train_processed)[1:30])

```




## word frequency



This step focus on how to choose optimal parameter for TF-IDF. We gonna plot the word frequency bar chart. The stopwords should be removed since meaningless, and also the low-frequency words for being noisy.

Since we have feature selection step in the further, we would rather import more words here.

```{r}


# clean function
clean_words <- function(data, column_name) {
  
  # 1. convert to tidy (Tokenization & Cleaning)
  # unnest_tokens will transfer to lowercase and delete non-letter chars
  tidy_text <- data %>%
    select(all_of(column_name)) %>%
    unnest_tokens(word, !!sym(column_name))
  
  # total words
  wa <- tidy_text %>% distinct(word) %>% nrow()
  
  # 2. filter stopwords and low frequency
  # custom_stopwords is defined before
  cleaned_text <- tidy_text %>%
    filter(!word %in% custom_stopwords) %>%
    group_by(word) %>%
    filter(n() >= 5) %>% # drop words less than 5 times
    ungroup()
  
  # word number after filter
  wb <- cleaned_text %>% distinct(word) %>% nrow()
  
  # print the statistics
  cat(sprintf("%s original words: %d  stopword and lowfreq removed: %d\n", 
              column_name, wa, wb))
  
  # 3. calculate the top 20
  word_counts <- cleaned_text %>%
    count(word, sort = TRUE) %>%
    slice_head(n = 20)
  
  # 4. visualization
  p <- ggplot(word_counts, aes(x = reorder(word, -n), y = n)) +
    geom_col(fill = "steelblue") +
    labs(title = paste("Top Word Frequencies of", column_name),
         x = "Words", y = "Frequency") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  print(p)
  
  # return complete word frequency
  full_counts <- cleaned_text %>%
    count(word, sort = TRUE)
  
  return(full_counts)
}

results <- list()
for (col in cols) {
  results[[col]] <- clean_words(df, col)
}
```


# Quit Model

This model focus on whether people quit their job, and what factor affect mostly in their decision of job-quitting.

Here we use Lasso-Logistics to select most important factor, Ridge-Logistics, and also Elastic-Logistics.



## Lasso

```{r}

tab <- table(yq_train)
weights <- ifelse(yq_train == 1, 
                  sum(tab) / (2 * tab["1"]), 
                  sum(tab) / (2 * tab["0"]))


lambda <- 10^seq(1,-2,length = 100)
model.quit.lasso <- cv.glmnet(X_train_processed, yq_train, family = binomial, alpha = 1, lambda = lambda, weights = weights)

model.quit.lasso_coef <- predict(model.quit.lasso, s = model.quit.lasso$lambda.min, type = "coefficients")

model.quit.lasso_coefs <- as.matrix(model.quit.lasso_coef)
model.quit.lasso_vars <- data.frame(
  Variable = rownames(model.quit.lasso_coefs),
  Coefficient = as.numeric(model.quit.lasso_coefs)
) %>% filter(Coefficient != 0 & Variable != "(Intercept)") %>%
  arrange(Coefficient)


model.quit.lasso_vars

roc_obj <- roc(yq_train, predict(model.quit.lasso, newx = X_train_processed, type = "response"))
best_threshold.lasso <- coords(roc_obj, "best", ret = "threshold")[1,1]

yq_pred.lasso <- ifelse(predict(model.quit.lasso, newx = X_test_processed, type = "response")>best_threshold.lasso, 1, 0)
log_roc_curve.lasso <- roc(yq_test, yq_pred.lasso)
plot(log_roc_curve.lasso)

auc(log_roc_curve.lasso)
```


## Ridge


```{r}
model.quit.ridge <- cv.glmnet(X_train_processed, yq_train, family = binomial, alpha = 0, lambda = lambda, weights = weights)

model.quit.ridge_coef <- predict(model.quit.ridge, s = model.quit.ridge$lambda.min, type = "coefficients")

model.quit.ridge_coefs <- as.matrix(model.quit.ridge_coef)
model.quit.ridge_vars <- data.frame(
  Variable = rownames(model.quit.ridge_coefs),
  Coefficient = as.numeric(model.quit.ridge_coefs)
) %>% filter(Coefficient != 0 & Variable != "(Intercept)") %>%
  arrange(Coefficient)


model.quit.ridge_vars



roc_obj <- roc(yq_train, predict(model.quit.ridge, newx = X_train_processed, type = "response"))
best_threshold.ridge <- coords(roc_obj, "best", ret = "threshold")[1,1]

yq_pred.ridge <- ifelse(predict(model.quit.ridge, newx = X_test_processed, type = "response")>best_threshold.ridge, 1, 0)
log_roc_curve.ridge <- roc(yq_test, yq_pred.ridge)
plot(log_roc_curve.ridge)

auc(log_roc_curve.ridge)
```




## Elastic Net


```{r}
results = list()
for(a in seq(0,1,0.05)){
  model.quit.elastic <- cv.glmnet(X_train_processed, yq_train, family = binomial, alpha = a, lambda = lambda, weights = weights)
  results[[as.character(a)]] <- min(model.quit.elastic$cvm)
}

# find best alpha
best_alpha <- names(which.min(unlist(results)))
print(paste("The best Alpha is:", best_alpha))

model.quit.elastic <- cv.glmnet(X_train_processed, yq_train, family = binomial, alpha = best_alpha, lambda = lambda)

model.quit.elastic_coef <- predict(model.quit.elastic, s = model.quit.elastic$lambda.min, type = "coefficients")

model.quit.elastic_coefs <- as.matrix(model.quit.elastic_coef)
model.quit.elastic_vars <- data.frame(
  Variable = rownames(model.quit.elastic_coefs),
  Coefficient = as.numeric(model.quit.elastic_coefs)
) %>% filter(Coefficient != 0 & Variable != "(Intercept)") %>%
  arrange(Coefficient)


model.quit.elastic_vars



roc_obj <- roc(yq_train, predict(model.quit.elastic, newx = X_train_processed, type = "response"))
best_threshold.elastic <- coords(roc_obj, "best", ret = "threshold")[1,1]

yq_pred.elastic <- ifelse(predict(model.quit.elastic, newx = X_test_processed, type = "response")>best_threshold.elastic, 1, 0)
log_roc_curve.elastic <- roc(yq_test, yq_pred.elastic)
plot(log_roc_curve.elastic)

auc(log_roc_curve.elastic)
```



## Report

```{r}
get_metrics_bin <- function(y_true, y_pred, name,
    positive = "1") {
  cm_obj <- confusionMatrix(
    data = as.factor(y_pred),
    reference = as.factor(y_true),
    positive = positive
  )
  out <- c(
    Sensitivity = unname(cm_obj$byClass["Sensitivity"]),
    Specificity = unname(cm_obj$byClass["Specificity"]),
    Accuracy    = unname(cm_obj$overall["Accuracy"]),
    AOC = auc(roc(yq_test, yq_pred.ridge))
  )
  return (c(Name = name, round(out, 4)))
}


rbind(
  get_metrics_bin(yq_test, yq_pred.lasso, "Lasso-Logistics"),
  get_metrics_bin(yq_test, yq_pred.ridge, "Ridge-Logistics"),
  get_metrics_bin(yq_test, yq_pred.elastic, "Elastic-Logistics")

)


```

# Length Model

Employee length is a fat tail variable and extremely right-skewed. Box-Cox trsnaformation applied here to deal with this problem.

```{r}
hist(yl_train, freq = F)

p_load(forecast)

lambda <- BoxCox.lambda(yl_train)

print(paste("Optimal Lambda:", lambda))

yl_train2 <- BoxCox(yl_train, lambda)
yl_test2 <- BoxCox(yl_test, lambda)


hist(yl_train2, freq = F)

```



## Lasso

```{r}
lambda <- 10^seq(1,-2,length = 100)
model.length.lasso <- cv.glmnet(X_train_processed, yl_train, alpha = 1, lambda = lambda)

model.length.lasso_coef <- predict(model.length.lasso, s = model.length.lasso$lambda.min, type = "coefficients")

model.length.lasso_coefs <- as.matrix(model.length.lasso_coef)
model.length.lasso_vars <- data.frame(
  Variable = rownames(model.length.lasso_coefs),
  Coefficient = as.numeric(model.length.lasso_coefs)
) %>% filter(Coefficient != 0 & Variable != "(Intercept)") %>%
  arrange(Coefficient)


model.length.lasso_vars

```


## Ridge


```{r}
model.length.ridge <- cv.glmnet(X_train_processed, yq_train, family = binomial, alpha = 0, lambda = lambda)

model.length.ridge_coef <- predict(model.length.ridge, s = model.length.ridge$lambda.min, type = "coefficients")

model.length.ridge_coefs <- as.matrix(model.length.ridge_coef)
model.length.ridge_vars <- data.frame(
  Variable = rownames(model.length.ridge_coefs),
  Coefficient = as.numeric(model.length.ridge_coefs)
) %>% filter(Coefficient != 0 & Variable != "(Intercept)") %>%
  arrange(Coefficient)


model.length.ridge_vars
```




## Elastic Net


```{r}
results = list()
for(a in seq(0,1,0.05)){
  model.length.elastic <- cv.glmnet(X_train_processed, yq_train, family = binomial, alpha = a, lambda = lambda)
  results[[as.character(a)]] <- min(model.length.elastic$cvm)
}

# find best alpha
best_alpha <- names(which.min(unlist(results)))
print(paste("The best Alpha is:", best_alpha))

model.length.elastic <- cv.glmnet(X_train_processed, yq_train, family = binomial, alpha = best_alpha, lambda = lambda)

model.length.elastic_coef <- predict(model.length.elastic, s = model.length.elastic$lambda.min, type = "coefficients")

model.length.elastic_coefs <- as.matrix(model.length.elastic_coef)
model.length.elastic_vars <- data.frame(
  Variable = rownames(model.length.elastic_coefs),
  Coefficient = as.numeric(model.length.elastic_coefs)
) %>% filter(Coefficient != 0 & Variable != "(Intercept)") %>%
  arrange(Coefficient)


model.quit.elastic_vars
```


