---
title: "EDA_Brian"
author: "Your Name"
date: "2026-04-09"
output:
  pdf_document: default
  html_document: default
---

# Loading Libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(dplyr)
library(corrplot)
library(here)
library(scales)
library(forcats)
```

# Read from current data version
```{r}
df <- read.csv(here("data", "processed", "glassdoor_reviews_cleaned_v2.csv"))

dim(df)
head(df)
```

# To make plots more appealing
```{r}
theme_set(
  theme_minimal(base_size = 13) +
    theme(
      plot.title = element_text(face = "bold", size = 15),
      plot.subtitle = element_text(size = 11),
      axis.title = element_text(face = "bold"),
      panel.grid.minor = element_blank(),
      legend.position = "bottom"
    )
)
```

# 1. Missing Data Analysis
```{r}
colSums(is.na(df))
```
### Explanation:
The variable employee_job_end_year contains missing values that correspond to
current employees who have not left their jobs. These are structurally missing
values and were intentionally preserved.

# 2. Overall Rating Distribution
```{r}
ggplot(df, aes(x = factor(rating_overall))) +
  geom_bar(fill = "steelblue", width = 0.7) +
  labs(
    title = "Distribution of Overall Ratings",
    subtitle = "Employee reviews are concentrated at lower ratings",
    x = "Overall Rating",
    y = "Number of Reviews"
  )
```

### Explanation: 
The distribution of overall ratings is skewed toward lower values, with a large
number of reviews concentrated around ratings of 1 and 2. This indicates that 
negative reviews are more common in this dataset. Higher ratings (4–5) appear 
less frequently, suggesting that employees are more likely to report negative experiences than positive ones.

# 3. Employee Length Distribution
```{r}
ggplot(df, aes(x = employee_length)) +
  geom_histogram(binwidth = 1, fill = "darkgreen", color = "white") +
  scale_x_continuous(breaks = pretty_breaks()) +
  labs(
    title = "Distribution of Employee Tenure",
    subtitle = "Most reviewers have relatively short tenures",
    x = "Employee Length (Years)",
    y = "Number of Reviews"
  )
```

### Explanation:
The distribution of employee tenure is right-skewed, with most employees having relatively short lengths of employment, typically between 1 and 3 years. As
tenure increases, the number of employees decreases significantly. A small 
number of employees have long tenures, which appear as outliers in the 
distribution. This pattern suggests that employee turnover may be relatively 
high, with fewer individuals remaining at a company for extended periods.

# 4. Recommendation Distribution
```{r}
ggplot(df, aes(x = flags_recommend_frend, fill = flags_recommend_frend)) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c("NEGATIVE" = "#C44E52", "POSITIVE" = "#4C72B0")) +
  labs(
    title = "Distribution of Recommendation Outcomes",
    subtitle = "Negative recommendations are more common in this dataset",
    x = "Recommendation",
    y = "Number of Reviews"
  )
```

### Explanation
The distribution of recommendation outcomes shows a clear imbalance, with 
significantly more negative recommendations than positive ones. This indicates 
that a larger proportion of employees in the dataset do not recommend their 
company. The imbalance suggests that negative experiences may be more frequently reported, which is consistent with the tendency for dissatisfied individuals to 
leave reviews more often than satisfied ones.

# 5. Overall Rating vs Recommendation
```{r}
ggplot(df, aes(x = flags_recommend_frend, y = rating_overall, fill = flags_recommend_frend)) +
  geom_boxplot(alpha = 0.8, show.legend = FALSE) +
  scale_fill_manual(values = c("NEGATIVE" = "#C44E52", "POSITIVE" = "#4C72B0")) +
  labs(
    title = "Overall Rating by Recommendation Outcome",
    subtitle = "Employees who recommend their company give substantially higher ratings",
    x = "Recommendation",
    y = "Overall Rating"
  )
```

### Explanation
There is a strong relationship between overall rating and recommendation 
behavior. Reviews with positive recommendations tend to have much higher overall ratings, while negative recommendations are associated with lower ratings. This suggests that overall rating is one of the strongest indicators of whether an
employee would recommend the company.

# 6. Correlation Between Rating Variables
```{r}
ratings <- df %>%
  select(
    rating_overall,
    rating_work_life,
    rating_culture_values,
    rating_compensation_benefits,
    rating_senior_leadership,
    rating_career_opportunities
  )

corr_matrix <- cor(ratings, use = "complete.obs")

corrplot(
  corr_matrix,
  method = "color",
  type = "upper",
  addCoef.col = "black",
  tl.col = "black",
  tl.srt = 30,
  number.cex = 0.7
)
```

### Explanation
The correlation matrix shows that the workplace rating variables are positively
related to one another. In particular, overall rating tends to move together
with culture, leadership, and career opportunities. This means that when 
employees rate one important workplace dimension highly, they often rate other dimensions highly as well, suggesting overlap among these measures.

# 7. Ratings by Sector
```{r}
ggplot(df, aes(x = fct_reorder(sector, rating_overall, .fun = median), y = rating_overall)) +
  geom_boxplot(fill = "skyblue", alpha = 0.8, outlier.alpha = 0.4) +
  coord_flip() +
  labs(
    title = "Overall Rating by Sector",
    subtitle = "Median ratings vary noticeably across industries",
    x = "Sector",
    y = "Overall Rating"
  )
```

### Explanation
Employee ratings vary across sectors, with some industries showing higher median satisfaction than others. The spread within each sector also differs, meaning 
that some sectors have more consistent employee experiences while others are 
more mixed. This suggests that industry context plays an important role in 
shaping employee satisfaction.

# 8. Recommendation Rate by Sector
```{r}
ggplot(df, aes(x = fct_infreq(sector), fill = flags_recommend_frend)) +
  geom_bar(position = "fill") +
  coord_flip() +
  scale_y_continuous(labels = percent_format()) +
  scale_fill_manual(values = c("NEGATIVE" = "#C44E52", "POSITIVE" = "#4C72B0")) +
  labs(
    title = "Recommendation Rate by Sector",
    subtitle = "Some industries have much higher shares of positive recommendations",
    x = "Sector",
    y = "Proportion of Reviews",
    fill = "Recommendation"
  )
```

### Explanation
The share of positive and negative recommendations differs across sectors. Some
sectors have relatively balanced outcomes, while others are dominated by 
negative recommendations. This reinforces the earlier finding that industry 
matters and suggests that workplace satisfaction is not evenly distributed 
across the economy.

# 9. Average Overall Rating By Year
```{r}
df$rating_date <- as.Date(df$rating_date)
df$rating_year <- as.numeric(format(df$rating_date, "%Y"))

yearly_avg <- df %>%
  group_by(rating_year) %>%
  summarise(avg_rating = mean(rating_overall, na.rm = TRUE), .groups = "drop")

ggplot(yearly_avg, aes(x = rating_year, y = avg_rating)) +
  geom_line(color = "#C44E52", linewidth = 1) +
  geom_point(color = "#C44E52", size = 2.5) +
  scale_x_continuous(breaks = yearly_avg$rating_year) +
  labs(
    title = "Average Overall Rating by Year",
    subtitle = "Employee ratings change over time rather than remaining constant",
    x = "Year",
    y = "Average Overall Rating"
  )
```

### Explanation
Average ratings show a noticeable time trend rather than staying constant across
years. The pattern suggests that employee sentiment changes over time, possibly reflecting shifts in workplace policies, labor-market conditions, or broader 
economic and social events.

# 10. COVID Impact Analysis
```{r}
ggplot(df, aes(x = factor(covid_related_review), y = rating_overall, fill = factor(covid_related_review))) +
  geom_boxplot(show.legend = FALSE, alpha = 0.8) +
  scale_fill_manual(values = c("0" = "#999999", "1" = "#4C72B0")) +
  scale_x_discrete(labels = c("0" = "Non-COVID", "1" = "COVID-Related")) +
  labs(
    title = "Overall Rating: COVID-Related vs Non-COVID Reviews",
    subtitle = "Pandemic-related reviews can be compared directly to the rest of the sample",
    x = "Review Type",
    y = "Overall Rating"
  )
```

### Explanation
Comparing COVID-related reviews with non-COVID reviews helps show whether the 
pandemic period was associated with different employee experiences. Both have a
median rating of approximately 2, indicating generally low employee satisfaction
across the dataset. This suggests that the pandemic did not significantly change
overall employee sentiment in this dataset, and that negative experiences were
already common before COVID.

# 11. Employee Length vs. Rating
```{r}
ggplot(df, aes(x = employee_length, y = rating_overall)) +
  geom_jitter(width = 0.2, height = 0.15, alpha = 0.25, color = "steelblue") +
  geom_smooth(method = "lm", se = FALSE, color = "darkred", linewidth = 1) +
  labs(
    title = "Employee Tenure vs Overall Rating",
    subtitle = "The relationship appears weak overall",
    x = "Employee Length (Years)",
    y = "Overall Rating"
  )
```

### Explanation
The relationship between tenure and overall rating appears weak. Although there
may be slight changes across different lengths of employment, employee 
satisfaction does not seem to be strongly determined by how long someone stayed
at the company. This suggests that other factors, such as culture, management, 
or compensation, may matter more than tenure alone.

# 12. Top Companies By Rating
```{r}
top_companies <- df %>%
  group_by(company_name) %>%
  summarise(
    avg_rating = mean(rating_overall, na.rm = TRUE),
    n_reviews = n(),
    .groups = "drop"
  ) %>%
  filter(n_reviews >= 5) %>%
  arrange(desc(avg_rating)) %>%
  slice_head(n = 10)

ggplot(top_companies,
       aes(x = reorder(company_name, avg_rating), y = avg_rating)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Top Companies by Average Rating",
    subtitle = "Among companies with at least 5 reviews",
    x = "Company",
    y = "Average Rating"
  )
```

### Explanation
The highest-rated companies in the filtered chart represent organizations with
both strong average ratings and a minimum number of reviews. Requiring multiple
reviews makes the comparison more reliable, since companies with only one or two reviews can appear artificially strong. This plot highlights companies that 
stand out more credibly in employee satisfaction.

# 13. Boxplot On Ratings By Employee Status
```{r}
status_plot_df <- df %>%
  mutate(
    employee_status = trimws(as.character(employee_status)),
    employee_status = ifelse(employee_status == "" | is.na(employee_status),
                             NA, employee_status)
  ) %>%
  filter(!is.na(employee_status))

ggplot(status_plot_df,
       aes(x = fct_reorder(employee_status, rating_overall, .fun = median),
           y = rating_overall)) +
  geom_boxplot(fill = "skyblue", alpha = 0.85, outlier.alpha = 0.35) +
  coord_flip() +
  labs(
    title = "Overall Rating by Employee Status",
    subtitle = "Blank or missing employee-status values were removed",
    x = "Employee Status",
    y = "Overall Rating"
  )
```

### Explanation
Interns report slightly higher satisfaction, while contract employees report 
somewhat lower ratings. However, the large overlap across groups suggests that 
employee status is not a major factor in determining overall ratings. The 
freelance category has very few observations and is not reliable.
