---
title: "AMS597 Project Data Cleaning"
author: "Group 1"
date: "2026-03-12"
output: 
  pdf_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries
```{r}
library(dplyr)
library(stringr)
library(lubridate)
library(tidyr)
```

# Reading the File
```{r}
library(here)

getwd()

list.files()

df <- read.csv("glassdoor_reviews_with_sector_revised.csv")

colSums(is.na(df))
```

# Removing duplicate reviews
```{r}
df_clean <- df %>%
  distinct(review_id, .keep_all = TRUE)

sum(duplicated(df_clean$review_id))
```

# Clean date column
```{r}
df_clean <- df_clean %>%
  mutate(
    rating_date = gsub('"', "", rating_date),
    rating_date = as.POSIXct(rating_date, format = "%Y-%m-%dT%H:%M:%OSZ", tz = "UTC"),
    rating_date = as.Date(rating_date),
    rating_year = year(rating_date)
  )

# Converts the timestamp into a proper date format for time-based analysis.
```

# Fixing helpful / unhelpful counts
```{r}
df_clean <- df_clean %>%
  mutate(
    count_helpful = as.numeric(ifelse(is.na(count_helpful), 0, count_helpful)),
    count_unhelpful = as.numeric(ifelse(is.na(count_unhelpful), 0, count_unhelpful))
  ) %>%
  mutate(
    helpfulness_score = count_helpful-count_unhelpful
  )

colSums(is.na(df_clean))
# Filling missing locations to unknown and standardizes
```

# Rename and clean COVID flag
```{r}
df_clean <- df_clean %>%
  rename(covid_related_review = flag_covid) %>%
  mutate(
    covid_related_review = ifelse(
      is.na(covid_related_review),
      ifelse(rating_year >= 2020 & rating_year <= 2023, 1, 0),
      ifelse(covid_related_review == TRUE, 1, 0)
    )
  )

# Renaming the column and converting to binary
```

# Converting numeric columns
```{r}
df_clean <- df_clean %>%
  mutate(
    employee_job_end_year = as.numeric(employee_job_end_year),
    employee_length = as.numeric(employee_length),
  ) %>% 
  mutate(
    rating_culture_values = as.numeric(rating_culture_values),
    rating_diversity_inclusion = as.numeric(rating_diversity_inclusion),
    rating_overall = as.numeric(rating_overall),
    rating_senior_leadership = as.numeric(rating_senior_leadership),
    rating_career_opportunities = as.numeric(rating_career_opportunities)
  )

```

# Cleaning Employee Type
```{r}
df_clean <- df_clean %>%
  mutate(
    employee_type = ifelse(
      is.na(employee_type) & !is.na(employee_job_end_year),
      "Former Employee",
      employee_type
    ),
    employee_type = ifelse(
      is.na(employee_type) & is.na(employee_job_end_year),
      "Current Employee",
      employee_type
    )
  )
```

# Impute rating_diversity_inclusion by company median
```{r}
df_clean <- df_clean %>%
  group_by(company_name) %>%
  mutate(
    rating_diversity_inclusion = ifelse(
      is.na(rating_diversity_inclusion),
      median(rating_diversity_inclusion, na.rm = TRUE),
      rating_diversity_inclusion
    )
  ) %>%
  ungroup()
```

```{r}
sum(is.na(df_clean$rating_diversity_inclusion))
```

# Sector-Level Fallback
```{r}
df_clean <- df_clean %>%
  group_by(sector) %>%
  mutate(
    rating_diversity_inclusion = ifelse(
      is.na(rating_diversity_inclusion),
      median(rating_diversity_inclusion, na.rm = TRUE),
      rating_diversity_inclusion
    )
  ) %>%
  ungroup()
```

```{r}
sum(is.na(df_clean$rating_diversity_inclusion))
```

# Median imputation for other numeric columns
```{r}
#calculate the % of missing data of rating_career_opportunities
mean(is.na(df_clean$rating_career_opportunities)) * 100

df_clean <- df_clean %>%
  mutate(
    employee_length = ifelse(
      is.na(employee_length),
      median(employee_length, na.rm = TRUE),
      employee_length
    ),
    rating_career_opportunities = ifelse(
      is.na(rating_career_opportunities),
      median(rating_career_opportunities, na.rm = TRUE),
      rating_career_opportunities
    )
  )
```

# Feature Engineering
```{r}
df_clean <- df_clean %>%
  mutate(
    employee_job_quit = grepl("^Former", employee_type, ignore.case = TRUE),
    employee_job_start_year = ifelse(
      employee_job_quit & !is.na(employee_job_end_year),
      employee_job_end_year - employee_length,
      rating_year - employee_length
    ),
    has_review_text = ifelse(
      (is.na(review_pros) | str_trim(review_pros) == "") &
      (is.na(review_cons) | str_trim(review_cons) == "") &
      (is.na(summary) | str_trim(summary) == ""),
      FALSE,
      TRUE
    )
  )
```

# Fill the reviews
```{r}
df_clean <- df_clean %>%
  mutate(
    summary = ifelse(is.na(summary) | str_trim(summary) == "", "", summary),
    advice_to_management = ifelse(is.na(advice_to_management) | str_trim(advice_to_management) == "", "", advice_to_management),
    review_pros = ifelse(is.na(review_pros) | str_trim(review_pros) == "", "", review_pros),
    review_cons = ifelse(is.na(review_cons) | str_trim(review_cons) == "", "", review_cons)
  )
```

# Dropping unecessary columns
```{r}
df_clean <- df_clean %>%
  select(
    -timestamp,
    -overview_id,
    -review_url,
    -employee_responses,
    -employee_type,
    -flag_featured,
    -review_advice,
    -url,
    -overview_url,
    -original_url,
    -glassdoor_employer_id
  )
```

# Final checks
```{r}
colSums(is.na(df_clean))
dim(df_clean)
```

# Saving cleaned file
```{r}
write.csv(
  df_clean,
  here("data", "processed", "glassdoor_reviews_cleaned.csv"),
  row.names = FALSE
)
```
