---
title: "AMS597 Project Data Cleaning"
author: "Group 1"
date: "2026-03-12"
output: 
  pdf_document:
    number_sections: true
---

```{r}
pkg_needed <- c("tidyverse", 
                "lubridate", 
                "ggplot2",
                "ggcorrplot",
                "here",
                "corrplot")
for (p in pkg_needed) {
  if (!requireNamespace(p, quietly = TRUE)) {
    install.packages(p)
  }
}

```


# Read Data
```{r}
library(here)
df <- read.csv(
  here("data", "processed", "chatgpt_glassdoor_reviews_with_sector_revised.csv")
)
head(df)
```


# Missing
```{r}
source("src/utils/data_helpers.R")

#Trimming whitespace in empty cells first
df <- Clean.Whitespace.NA(df)
Missing.Summary(df)
nrow(df)
```


# Process Employee Status

If ```employee_length``` is missing, then delete it. (4 rows)

If is current employee, then use rating_year instead of end_year to minus employee_length to calculate the start_year. After this step we make sure all missing of end_year are currently employees.

```{r}

library(tidyverse)
library(lubridate)

dropped_employee_length <- df %>% select(everything()) %>% filter(is.na(employee_length))
dropped_employee_length

df <- df %>% 
  drop_na(employee_length)

df$rating_year = as.numeric(sub("[^0-9]*(\\d{4}).*", "\\1", df$rating_date))

df[is.na(df$employee_type), "employee_type"] <- "Former"
# fill it as Former since they're all 
# todo  make it better: check whether endyear exists

df$employee_job_quit <- ifelse(grepl("^Former", df$employee_type, ignore.case = TRUE),TRUE, FALSE)

df$employee_job_start_year <- ifelse(
  df$employee_job_quit,
  df$employee_job_end_year - df$employee_length, # Former use the year quit
  df$rating_year - df$employee_length # Current: use the year of rating
)

lengthvars <- c("employee_job_start_year", "employee_job_end_year", "employee_job_quit", "employee_length", "employee_type")
print(df[, lengthvars])

# is there anyone quit job without giving end year? no
# all missing of employee_job_end_year due to not quit job
df[df$employee_job_quit & is.na(df$employee_job_end_year), ]


```
## Default missing count_helpful and count_unhelpful values to 0, get net_vote

```{r}
df <- df %>%
  mutate(
    count_helpful   = coalesce(count_helpful, 0),
    count_unhelpful = coalesce(count_unhelpful, 0),
    net_vote = count_helpful - count_unhelpful,
    flag_covid = coalesce(as.logical(flag_covid), FALSE)
  )
Missing.Summary(df)
nrow(df)
```

## Missingness Correlation Matrix
```{r}
missing_matrix <- is.na(df)
missing_matrix <- missing_matrix * 1 #coerces boolean to numeric T = 1, F = 0
missing_matrix <- missing_matrix[, colSums(missing_matrix) > 0] #filter out columns with no missing 
cor_matrix <- cor(missing_matrix)

library(ggcorrplot)
ggcorrplot(cor_matrix, tl.srt = 90, tl.cex = 8)

library(corrplot)
corrplot(cor_matrix, method = "color", tl.cex = 0.6)

```
```{r}
cor_long <- as.data.frame(as.table(cor_matrix))

cor_long <- cor_long[cor_long$Var1 != cor_long$Var2, ] # avoids listing same variables
cor_long <- cor_long %>%arrange(desc(abs(Freq)))%>%filter(!is.na(Freq)) #sorts descending, filters out NA correlation 

cor_long
```



## Removing Columns
```{r}
cols_to_remove <- c("employee_responses", 
                    "flag_featured",
                    "glassdoor_employer_id",
                    "url",
                    "original_url",
                    "overview_url",
                    "review_url",
                    "overview_id",
                    "count_helpful",
                    "count_unhelpful",
                    "glassdoor_employer_id",
                    "timestamp"
                    )
newdf <-Remove.Columns(df, cols_to_remove)
names(newdf)
Missing.Summary(newdf)
```
## Recorrelate
```{r}
missing_matrix2 <- is.na(newdf)
missing_matrix2 <- missing_matrix2 * 1 #coerces boolean to numeric T = 1, F = 0
missing_matrix2 <- missing_matrix2[, colSums(missing_matrix2) > 0] #filter out columns with no missing 
cor_matrix2 <- cor(missing_matrix2)
cor_new <- as.data.frame(as.table(cor_matrix2))

cor_new <- cor_new[cor_new$Var1 != cor_new$Var2, ] # avoids listing same variables
cor_new <- cor_new %>%arrange(desc(abs(Freq)))%>%filter(!is.na(Freq)) #sorts descending, filters out NA correlation 

library(ggcorrplot)
ggcorrplot(cor_matrix2, tl.srt = 90, tl.cex = 8)

cor_new
```

```{r}
# write to csv
write.csv(
  newdf,
  here("data", "processed", "data_cleaned_mahin-ver.csv"),
  row.names = FALSE
)
```



