---
title: "AMS597 Project Data Cleaning"
author: "Group 1"
date: "2026-03-12"
output: 
  pdf_document:
    number_sections: true
---

```{r}
pkg_needed <- c("tidyverse", 
                "lubridate", 
                "ggplot2",
                "ggcorrplot",
                "here",
                "naniar")
for (p in pkg_needed) {
  if (!requireNamespace(p, quietly = TRUE)) {
    install.packages(p)
  }
}

```


# Read Data
```{r}
library(here)
df <- read.csv(
  here("data", "processed", "chatgpt_glassdoor_reviews_with_sector_revised.csv")
)
head(df)
```


# Missing
```{r}
source("src/utils/data_helpers.R")

#Trimming whitespace in empty cells first
Clean.Whitespace.NA(df)

Missing.Check.By.Column(df)
```


# Process Employee Status

If ```employee_length``` is missing, then delete it. (4 rows)

If is current employee, then use rating_year instead of end_year to minus employee_length to calculate the start_year. After this step we make sure all missing of end_year are currently employees.

```{r}

library(tidyverse)
library(lubridate)

df <- df %>% 
  drop_na(employee_length)

df$rating_year = as.numeric(sub("[^0-9]*(\\d{4}).*", "\\1", df$rating_date))

df[is.na(df$employee_type), "employee_type"] <- "Former"
# fill it as Former since they're all 
# todo  make it better: check whether endyear exists

df$employee_job_quit <- ifelse(grepl("^Former", df$employee_type, ignore.case = TRUE),TRUE, FALSE)

df$employee_job_start_year <- ifelse(
  df$employee_job_quit,
  df$employee_job_end_year - df$employee_length, # Former use the year quit
  df$rating_year - df$employee_length # Current: use the year of rating
)

lengthvars <- c("employee_job_start_year", "employee_job_end_year", "employee_job_quit", "employee_length", "employee_type")
print(df[, lengthvars])

# is there anyone quit job without giving end year? no
# all missing of employee_job_end_year due to not quit job
df[df$employee_job_quit & is.na(df$employee_job_end_year), ]

sort(Missing.Check.By.Column(df), descending = T)
```
## Default missing count_helpful and count_unhelpful values to 0, get net_vote

```{r}
df <- df %>%
  mutate(
    count_helpful   = coalesce(count_helpful, 0),
    count_unhelpful = coalesce(count_unhelpful, 0),
    net_vote = count_helpful - count_unhelpful,
    flag_covid = coalesce(as.logical(flag_covid), FALSE)
  )
Missing.
```

## Missingness Correlation Matrix
```{r}


library(ggcorrplot)

ggcorrplot(cor_matrix)

```



## Removing Columns
```{r}
cols_to_remove <- c("employee_responses", 
                    "flag_featured",
                    "glassdoor_employer_id",
                    "url",,
                    "original url",
                    
                    )
```

