---
title: "Dataset_demo"
author: "Quynh Tran"
date: "2026-02-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Import libraries
```{r}
library(MASS)
library(tidyverse)
#install.packages("lattice")
library(lattice)
library(caret)
#install.packages("Matrix")
library(Matrix)
library(glmnet)
library(mlbench)
library(caret)
```

#Import the dataset
```{r}
df <- read.csv("data/company_tagger/glassdoor_reviews_with_sector_revised.csv")
#view(df)
summary(df)
```

# EDA: 

# Some EDA
```{r}
#install.packages("skimr")
#library(skimr)
#skim(df)
```

```{r}
#install.packages("DataExplorer")
#library(DataExplorer)
#DataExplorer::create_report(df)
```

# Data Viz (before data cleaning)

# Visualizing patterns of missingness across variables.

# checking which variables tend to be missing together using a correlation heatmap of missing values.
```{r}
missing_matrix <- is.na(df)
missing_matrix <- missing_matrix * 1 #coerces boolean to numeric T = 1, F = 0
missing_matrix <- missing_matrix[, colSums(missing_matrix) > 0] #filter out columns with no missing 
cor_matrix <- cor(missing_matrix)

#install.packages("ggcorrplot")
library(ggcorrplot)
ggcorrplot(cor_matrix, tl.srt = 90, tl.cex = 8)

#install.packages("corrplot")
library(corrplot)
corrplot(cor_matrix, method = "color", tl.cex = 0.6)

```
```{r}
# which variables have the strongest missingness relationships
cor_long <- as.data.frame(as.table(cor_matrix))

cor_long <- cor_long[cor_long$Var1 != cor_long$Var2, ] # avoids listing same variables
cor_long <- cor_long %>%arrange(desc(abs(Freq)))%>%filter(!is.na(Freq)) #sorts descending, filters out NA correlation 

cor_long
```

# Data Cleaning:

```{r}
library(tidyverse)
library(lubridate)

df$employee_job_quit <- ifelse(grepl("^Former", df$employee_type, ignore.case = TRUE),TRUE, FALSE)

df$employee_job_start_year <- ifelse(
  df$employee_job_quit,
  df$employee_job_end_year - df$employee_length, # Former use the year quit
  df$rating_year - df$employee_length # Current: use the year of rating
)

dim(df)
#lengthvars <- c("employee_job_start_year", "employee_job_end_year", "employee_job_quit", "employee_length", "employee_type")
#print(df[, lengthvars])

# is there anyone quit job without giving end year? no
# all missing of employee_job_end_year due to not quit job
#df[df$employee_job_quit & is.na(df$employee_job_end_year), ]

```

# Create a column for employee_type1 that identifies whether an employee if former or current, based on column employee_type, by extracting the first two words from employee_type
```{r}
df$employee_type1 <- sapply(strsplit(df$employee_type, " "), `[`, 1)
#df$employee_type1
```

# For column count_helpful and count_unhelpful, replace missing values (NA) in count_helpful and count_unhelpful with 0
```{r}
df$count_helpful[is.na(df$count_helpful)] <- 0
df$count_unhelpful[is.na(df$count_unhelpful)] <- 0

# check if these columns still have any missing values
colSums(is.na(df[, c("count_helpful", "count_unhelpful")])) # passed!
```

# Create a new column called helpfulness score that stores the difference between count_helpful and count_unhelpful
```{r}
df$helpfulness_score <- df$count_helpful - df$count_unhelpful
df$helpfulness_score
```

# Replace missing values (NA) in the advice_to_management column with "No"
```{r}
df$advice_to_management[is.na(df$advice_to_management)] <- "No"
```

# For rating_diversity_inclusion, fill missing value using the MEDIAN within the same company. If the company has only one review (or no median available), then fill with the median of the sector.
```{r}
# Visualize distribution of rating_diversity_inclusion
library(ggplot2)

ggplot(df, aes(x = rating_diversity_inclusion)) +
  geom_histogram(bins = 20) +
  labs(
    title = "Distribution of Diversity & Inclusion Rating",
    x = "Rating",
    y = "Count"
  ) +
  theme_minimal()

# Step 1 — Define Median Safely (Handle All-NA Case)
safe_median <- function(x) {
  if (all(is.na(x))) return(NA)
  return(median(x, na.rm = TRUE))
}

sum(is.na(df$rating_diversity_inclusion))
# Step 2 — Company-Level Median Imputation
library(dplyr)

df <- df %>%
  group_by(company_name) %>%
  mutate(
    rating_diversity_inclusion = ifelse(
      is.na(rating_diversity_inclusion),
      safe_median(rating_diversity_inclusion),
      rating_diversity_inclusion
    )
  ) %>%
  ungroup()

# Step 3 — Sector-Level Fallback
df <- df %>%
  group_by(sector) %>%
  mutate(
    rating_diversity_inclusion = ifelse(
      is.na(rating_diversity_inclusion),
      safe_median(rating_diversity_inclusion),
      rating_diversity_inclusion
    )
  ) %>%
  ungroup()

# Step 4 — Final Check
sum(is.na(df$rating_diversity_inclusion)) # passed!

```
Diversity & Inclusion Rating data is ordinal (1-5 rating) and not normally distributed; hence, we will use median imputation for missing data. 

# Same median imputation for employee_length
```{r}
# Step 1: Visualize employee_length and check its distribution
# library(ggplot2)

ggplot(df, aes(x = employee_length)) +
  geom_histogram(bins = 30) +
  labs(
    title = "Distribution of Employee Length",
    x = "Employee Length (Years)",
    y = "Count"
  ) +
  theme_minimal()


sum(is.na(df$employee_length))
# Step 2 — Company-Level Median Imputation
library(dplyr)

df <- df %>%
  group_by(company_name) %>%
  mutate(
    employee_length = ifelse(
      is.na(employee_length),
      safe_median(employee_length),
      employee_length
    )
  ) %>%
  ungroup()

# Step 3 — Sector-Level Fallback
df <- df %>%
  group_by(sector) %>%
  mutate(
    employee_length = ifelse(
      is.na(employee_length),
      safe_median(employee_length),
      employee_length
    )
  ) %>%
  ungroup()

# Step 4 — Final Check
sum(is.na(df$employee_length)) # passed!

```
Employee length data is clearly right-skewed according to the histogram; hence, we're using median imputation for missing data for this column. 

# Remove certain columns that do not serve the analysis: overview_id, review_id, review_url, url, overview_url, original_url, flag_covid, review_advice, flag_featured: 
```{r}
df <- subset(df, select = -c(timestamp, overview_id, review_id, review_url, url, overview_url, original_url, employee_type, employee_responses, flag_covid, review_advice, flag_featured))
head(df)
dim(df) # after the removal, we have 1000 obs and 29 features. 
```

# Create useful variables: 
# Positive vs. Negative reviews: 
```{r}
df$sentiment <- ifelse(df$rating_overall >= 4, "positive",
                 ifelse(df$rating_overall <= 2, "negative", "neutral"))

head(df)
```

# Convert categorical variables:
In R, categorical variables should usually be converted to factors because factors tell R that the variable represents categories, not numeric values or free text. This affects how models, summaries, and plots treat the variable.
```{r}
# Employee-related categories
df$employee_status <- as.factor(df$employee_status)
df$employee_type1 <- as.factor(df$employee_type1)
df$employee_job_title <- as.factor(df$employee_job_title)
df$employee_location <- as.factor(df$employee_location)

# Recommendation / Flags
df$flags_recommend_frend <- as.factor(df$flags_recommend_frend)
df$flags_ceo_approval <- as.factor(df$flags_ceo_approval)
df$flags_business_outlook <- as.factor(df$flags_business_outlook)

# Company / Industry Variables
df$company_name <- as.factor(df$company_name)
df$sector <- as.factor(df$sector)

# Sentiment
df$sentiment <- as.factor(df$sentiment)
```

```{r}
str(df)
```

# Descriptive Statistics: 
```{r}
summary(df$rating_overall)
summary(df$rating_work_life)
summary(df$rating_culture_values)
summary(df$rating_career_opportunities)
summary(df$rating_compensation_benefits)
summary(df$rating_diversity_inclusion)
summary(df$rating_senior_leadership)
```

```{r}
mean(df$rating_overall)
sd(df$rating_overall)
```

# For rating_date column, extract the first 10 characters and convert them to date format: 
```{r}
# Step 1: Inspect the column first
sum(is.na(df$rating_date))
sum(df$rating_date == "", na.rm = TRUE)
head(df$rating_date)
unique(df$rating_date)[1:10]

# Step 2: Convert to character first, then extract the date:
df$rating_date_clean <- gsub('"', "", as.character(df$rating_date))
df$rating_date_clean <- as.Date(substr(df$rating_date_clean, 1, 10), format = "%Y-%m-%d")

# Step 3: Verify
str(df$rating_date_clean)
head(df$rating_date_clean)
```


# Data Viz:

# Histogram on distribution of rating_overall:
```{r}
library(ggplot2)

ggplot(df, aes(rating_overall)) +
  geom_histogram(binwidth=1, fill="steelblue")
```

# Heatmap on relationships between rating categories
```{r}
#install.packages("corrplot")
library(corrplot)

ratings <- df[, c(
  "rating_overall",
  "rating_work_life",
  "rating_culture_values",
  "rating_compensation_benefits",
  "rating_senior_leadership",
  "rating_career_opportunities"
)]

corr_matrix <- cor(ratings, use = "complete.obs")

corrplot(corr_matrix,
         method = "color",
         col = colorRampPalette(c("white", "red"))(200),
         type = "upper",
         addCoef.col = "black",
         tl.col = "black",
         tl.srt = 45)
```
Rating_overall is most correlated with Rating_culture_values, Rating_senior_leadership, and Rating_career_opportunities, and least correlated with Rating_work_life and Rating_compensation_benefits. 

# Boxplot on Ratings by Employee Status
```{r}
ggplot(df, aes(employee_status, rating_overall)) +
  geom_boxplot()
```

# Scatterplot with Trend Line of Rating Overall: 

```{r}
ggplot(df, aes(x = rating_date_clean, y = rating_overall)) +
  geom_point(alpha = 0.3, color = "steelblue") +
  geom_smooth(method = "loess", color = "red") +
  labs(
    title = "Overall Rating Over Time",
    x = "Review Date",
    y = "Overall Rating"
  )

# another plot for after 2020
```
```{r}
df$rating_year <- format(df$rating_date_clean, "%Y")

library(dplyr)

yearly_avg <- df %>%
  group_by(rating_year) %>%
  summarise(avg_rating = mean(rating_overall, na.rm = TRUE))

ggplot(yearly_avg, aes(x = rating_year, y = avg_rating, group = 1)) +
  geom_line(color = "darkred") +
  geom_point(size = 3) +
  labs(
    title = "Average Rating by Year",
    x = "Year",
    y = "Average Overall Rating"
  ) + ylim(1, 5)

# Same plot but we will only see from year 2020 and after
yearly_avg_filtered <- yearly_avg %>%
  filter(as.numeric(rating_year) > 2020)

ggplot(yearly_avg_filtered, aes(x = rating_year, y = avg_rating, group = 1)) +
  geom_line(color = "darkred") +
  geom_point(size = 3) +
  labs(
    title = "Average Rating by Year (After 2020)",
    x = "Year",
    y = "Average Overall Rating"
  ) +
  coord_cartesian(ylim = c(1, 5))
```
Observations from the line graph: average employee ratings decline from ~2015–2018 and then rise starting around 2019. 
Some plausible explanations: 
1️⃣ Changes in the Labor Market (Macroeconomic Effects)

Employee sentiment often correlates with economic cycles.

2015–2018

During strong labor markets:

unemployment fell steadily in many economies (e.g., U.S. unemployment dropped from ~5.3% in 2015 to ~3.9% in 2018)

employees had more outside options

When employees have better alternatives, they:

become more critical of employers

expect higher pay, benefits, flexibility

So reviews may become more negative.

After 2019

Several factors could cause ratings to increase:

companies improving benefits (remote work, flexibility)

economic uncertainty making employees more appreciative of stable jobs

workplace reforms

This is why researchers sometimes study:

“Does employee satisfaction decrease when the labor market is strong?”

2️⃣ Platform / User Behavior Effects (Selection Bias)

Glassdoor-type review platforms often experience changes in who leaves reviews.

Possible scenarios:

Early years

Mostly dissatisfied employees leave reviews.

Later years

Companies encourage employees to leave reviews.

Example:

HR campaigns

onboarding prompts

employer branding initiatives

This increases the share of positive reviews, which raises the average rating.

This is a classic review selection bias.

3️⃣ Changes in Company Policies

Around the late 2010s many companies introduced:

better work-life balance policies

diversity & inclusion initiatives

remote work flexibility

improved benefits packages

These reforms could improve employee sentiment starting around 2019.

4️⃣ Generational / Cultural Changes in Workplaces

Workplace expectations evolved rapidly after 2018:

Examples:

flexible work arrangements

mental health awareness

focus on culture

Companies adapting to these changes may see improving ratings.

# Number of reviews over time (by year)
```{r}
# Convert rating_year column to numeric
df$rating_year <- as.numeric(df$rating_year)

# Count number of reviews per year
reviews_per_year <- df %>%
  group_by(rating_year) %>%
  summarise(n_reviews = n())

# Plot it
ggplot(reviews_per_year, aes(x = factor(rating_year), y = n_reviews)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Number of Reviews by Year",
    x = "Year",
    y = "Number of Reviews"
  ) +
  theme_minimal()
```

# Heatmap showing patterns of average rating by sector over years
```{r}
# Compute the yearly average by sector
sector_year_avg <- df %>%
  group_by(rating_year, sector) %>%
  summarise(avg_rating = mean(rating_overall, na.rm = TRUE))

# Heatmap
ggplot(sector_year_avg, aes(x = rating_year, y = sector, fill = avg_rating)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "red") +
  labs(
    title = "Average Rating Heatmap by Sector and Year",
    x = "Year",
    y = "Sector"
  ) +
  theme_minimal()
```
Observations: 
- Some sectors, such as Finance, Energy/Mining, Government, Marketing, and Construction, show higher average ratings after 2020. Possible interpretations are workplace policy improvements, remote work flexibility, and post-pandemic changes in employee expectations. 
- Healthcare sector shows mixed trends - higher average ratings earlier from 2013 to 2015, and appears more moderate later. This could have been due to increased stress or workload in later years and workforce burnout effects for healthcare professionals post-COVID 19. 

# Modeling: 

# Question 1: Predict whether an employee will recommend the company to their friends

# Check the target variable flags_recommend_frend
```{r}
unique(df$flags_recommend_frend) # 2 levels: NEGATIVE and POSITIVE
levels(df$flags_recommend_frend)
str(df$flags_recommend_frend)
```
flags_recommend_frend column is a binary categorical variable, and is a factor -> ready for the modeling. 

# Model 1: Logistic Regression

```{r}
table(df$flags_recommend_frend, useNA = "ifany")

str(df$flags_recommend_frend)
```
confirms this is a binary classification problem, and the target variable is a factor with 2 levels "positive" and "negative". 

# Step 1: Select Predictor Variables and Remove Missing Data (because logistic regression cannot handle missing data)
```{r}
data <- df[, c(
"flags_recommend_frend", 
"employee_length",
"employee_type1",
"flags_business_outlook",
"flags_ceo_approval",
"rating_overall",
"rating_work_life",
"rating_culture_values",
"rating_compensation_benefits",
"rating_senior_leadership",
"rating_diversity_inclusion"
)]

data <- na.omit(data)

dim(data)

names(data)

# check class proportion before splitting
prop.table(table(data$flags_recommend_frend))
```


# Step 2: Split the dataset into 80% training and 20% testing
```{r}
library(caret)

set.seed(123)

train_index <- createDataPartition(
  data$flags_recommend_frend,
  p = 0.8,
  list = FALSE
)

train.data <- data[train_index, ]
test.data <- data[-train_index, ]

# Check class imbalance after splitting
prop.table(table(train.data$flags_recommend_frend))
prop.table(table(test.data$flags_recommend_frend))
```
The class proportions (approximately 70% negative and 30% positive) were preserved in both training and testing sets. This guarantees that the model is trained and evaluated on representative data, leading to reliable performance estimates.

# Step 3: Fit the logistic regression model
```{r}
set.seed(123)
model_logit <- glm(
  flags_recommend_frend ~.,
  data = train.data,
  family = binomial
)

summary(model_logit)
```
Interpretations of the logistic regression model summary: 
1. Coefficients: 
positive coefficient → increases probability of recommending
negative coefficient → decreases probability

2. p-values:
Per the model summary, flags_business_outlookPOSITIVE, rating_overall, flags_ceo_approvalDISAPPROVE (decreases the probability an employee recommends the company), rating_work_life, rating_culture_values, flags_business_outlookNEUTRAL are statistically significant predictors for whether a person recommends a company, 5% significance level.  

# Odds-ratio: Converts log-odds coefficients
```{r}
exp(coef(model_logit))
```
Interpretation:

Odds ratios show how a one-unit increase in a predictor multiplicatively changes the odds of the outcome. A one-point increase in overall rating makes employees 2.68 times more likely to recommend the company, or a one-point increase in flags_business_outlookPOSITIVE makes employees 9.16 times more likely to recommend the company.

# Step 4: Predict on test data
```{r}
log_pred_prob <- predict(model_logit, test.data, type = "response")
log_pred_prob
```
These are probabilities of recommending the company. 

# Step 5: Evaluate different classification thresholds to select the best one using F1 scores
# F1 = 2 * (precision * recall) / (precision + recall)
```{r}
# Step 5.1: Create a Sequence of Thresholds
thresholds <- seq(0.1, 1.0, by = 0.1)

# Step 5.2: Define a function to calculate F1 score
f1_score <- function(actual, predicted) {

  tp <- sum(actual == "POSITIVE" & predicted == "POSITIVE")
  fp <- sum(actual == "NEGATIVE" & predicted == "POSITIVE")
  fn <- sum(actual == "POSITIVE" & predicted == "NEGATIVE")

  precision <- ifelse((tp + fp) == 0, 0, tp / (tp + fp))
  recall <- ifelse((tp + fn) == 0, 0, tp / (tp + fn))

  if (precision + recall == 0) return(0)

  f1 <- 2 * precision * recall / (precision + recall)

  return(f1)
}

# Step 5.3: Loop Through Thresholds
f1_results <- data.frame(threshold = thresholds, F1 = NA)

for (i in seq_along(thresholds)) {
  
  t <- thresholds[i]
  
  pred_class <- ifelse(log_pred_prob > t, "POSITIVE", "NEGATIVE")
  
  f1_results$F1[i] <- f1_score(test.data$flags_recommend_frend, pred_class)
}

# Step 5.4: Print F1 scores
print(f1_results)

# Step 5.5: Choose the threshold with the best F1 score
best_threshold <- f1_results$threshold[which.max(f1_results$F1)]
best_f1 <- max(f1_results$F1)

best_threshold
best_f1
```
Threshold of 0.4 produces the highest F1 score of 0.85

# Step 6: Convert Probabilities to Class Predictions using threshold of 0.4
```{r}
log_pred_class <- ifelse(log_pred_prob > best_f1, "POSITIVE", "NEGATIVE")

# Convert to factor
log_pred_class <- as.factor(log_pred_class)
```

# Step 7: Evaluate Model Performance
```{r}
logistic_cm <- confusionMatrix(log_pred_class, test.data$flags_recommend_frend)
logistic_cm
```
Logistic regression has Accuracy of 0.8776

# Step 8: ROC Curve and AUC
```{r}
library(pROC)

log_roc_curve <- roc(test.data$flags_recommend_frend, log_pred_prob)
plot(log_roc_curve)

auc(log_roc_curve)
```
Logistic Regression has AUC of 0.9403.

# Model 2: Decision Tree: 

# Step 1 — Load Required Packages for Decision Tree Classification
```{r}
# ---- 0. Install / load packages ----
pkg_needed <- c("caret", "rpart", "rattle")
for (p in pkg_needed) {
  if (!requireNamespace(p, quietly = TRUE)) {
    install.packages(p)
  }
}

#install.packages("bitops")
#install.packages("ggplot2")
#install.packages("lattice")
#install.packages("Matrix")
#install.packages("rpart.plot")
library(caret)
library(rpart)
library(rattle)
library(rpart.plot)
```

# Train the Decision Tree
```{r}
# Fit the initial tree
set.seed(123)
tree_model <- rpart(
  flags_recommend_frend ~ .,
  data = train.data,
  method = "class", 
  control = rpart.control(cp = 0, xval = 10, minsplit = 2, minbucket = 1)
)

# Visualize the tree
rpart.plot(tree_model, main = "Initial Tree")

# Check CP table
printcp(tree_model)

# Select the best CP value
best_cp <- tree_model$cptable[
  which.min(tree_model$cptable[, "xerror"]),
  "CP"
]

cat("\nBest CP value:", best_cp, "\n")

# Prune the tree using best CP value
pruned_tree <- prune(tree_model, cp = best_cp)

# Visualize the pruned tree
fancyRpartPlot(pruned_tree, main = "Optimal Pruned Tree")

# Predict on test data
pred_test <- predict(pruned_tree, newdata = test.data, type = "class")

# Confusion Matrix and Metrics
pruned_tree_cm <- confusionMatrix(pred_test, test.data$flags_recommend_frend)
pruned_tree_cm

```
Observations: 
- Overall rating is the dominant driver of recommendation
- rating overall and flags business outlook are the strongest variables to make good predictions.
- Many rating variables are likely highly correlated with rating_overall, so the tree doesn’t need them.

# ROC curve and AUC for pruned tree
```{r}
# ROC and AUC for pruned tree
#install.packages("pROC")
library(pROC)

# Make sure outcome levels are ordered correctly
test.data$flags_recommend_frend <- factor(
  test.data$flags_recommend_frend,
  levels = c("NEGATIVE", "POSITIVE")
)

# Predicted probabilities
pred_prob <- predict(pruned_tree, newdata = test.data, type = "prob")

# Probability of positive class
pred_prob_positive <- pred_prob[, "POSITIVE"]

# ROC curve
roc_pruned_tree <- roc(
  response = test.data$flags_recommend_frend,
  predictor = pred_prob_positive,
  levels = c("NEGATIVE", "POSITIVE"),
  direction = "<"
)

# AUC
auc_pruned_tree <- auc(roc_pruned_tree)
auc_pruned_tree

# Plot ROC
plot(
  roc_pruned_tree,
  col = "blue",
  lwd = 2,
  main = "ROC Curve - Pruned Decision Tree"
)

# Add diagonal reference line
abline(a = 0, b = 1, lty = 2, col = "gray")
```
Pruned Decision tree has AUC of 0.9056

# Random forest: 

# Step 1: Import Random Forest package
```{r}
#install.packages("randomForest")
#install.packages("ggplot2")
library(randomForest)
library(caret)
```

# Step 5: Fit the random forest model
```{r}
# Train Random Forest with 10-fold CV
set.seed(123)

p <- ncol(train.data) - 1

rf_model <- train(
  flags_recommend_frend ~ .,
  data = train.data,
  method = "rf",
  trControl = trainControl(method = "cv", number = 10),
  tuneGrid = expand.grid(mtry = 1:p),
  ntree = 500,
  importance = TRUE
)

# Find the best mtry
bestTune <- rf_model$bestTune
bestTune

# Refit the model with the best mtry
rf_model_tuned <- train(
  flags_recommend_frend ~ .,
  data = train.data,
  method = "rf",
  trControl = trainControl(
    method = "cv",
    number = 10,
    classProbs = TRUE,
    summaryFunction = twoClassSummary
  ),
  tuneGrid = expand.grid(mtry = bestTune),  # 👈 use your chosen mtry
  ntree = 500,
  metric = "ROC",
  importance = TRUE
)

# Variable importance plots
varImpPlot(rf_model_tuned$finalModel, type = 1, main = "Mean Decrease Accuracy (MSE)")  # Mean Decrease Accuracy (MSE)
varImpPlot(rf_model_tuned$finalModel, type = 2, main = "Mean Decrease Node Impurity")  # Mean Decrease Node Impurity

# Predict on test data
pred_test <- predict(rf_model_tuned, test.data)

# Confusion matrix and metrics
cm <- confusionMatrix(
  pred_test,
  test.data$flags_recommend_frend
)

cm

# Metrics
sensitivity <- cm$byClass["Sensitivity"]
specificity <- cm$byClass["Specificity"]
accuracy    <- cm$overall["Accuracy"]

sensitivity
specificity
accuracy

```

# Step 10: ROC curve and AUC
```{r}
library(pROC)
# Predict probabilities
rf_pred_prob <- predict(rf_model_tuned, newdata = test.data, type = "prob")

# Extract probability of POSITIVE class
rf_pred_prob_pos <- rf_pred_prob[, "POSITIVE"]

# Make Sure Target is Factor
test.data$flags_recommend_frend <- factor(
  test.data$flags_recommend_frend,
  levels = c("NEGATIVE", "POSITIVE")
)

# Compute ROC Curve
roc_rf <- roc(
  response = test.data$flags_recommend_frend,
  predictor = rf_pred_prob_pos,
  levels = c("NEGATIVE", "POSITIVE"),
  direction = "<"
)

# AUC
auc_rf <- auc(roc_rf)
auc_rf

# Plot ROC curve
plot(
  roc_rf,
  col = "blue",
  lwd = 2,
  main = "ROC Curve - Random Forest"
)

abline(a = 0, b = 1, lty = 2, col = "gray")
```
Random forest model has AUC of 0.9309.

Due to class imbalance, decision tree predicts majority class better, which is shown by better accuracy but worse AUC compared to random forest. Accuracy is defined by a threshold and AUC is defined without threshold, so AUC is the better metric when we haven’t tuned the threshold, because of class imbalance. 


