---
title: "modeling_with_cleaned_data"
author: "Quynh Tran"
date: "2026-04-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import packages
```{r}
library(MASS)
library(tidyverse)
#install.packages("lattice")
library(lattice)
library(caret)
#install.packages("Matrix")
library(Matrix)
library(glmnet)
library(mlbench)
library(caret)
library(here)
```

# Import dataset
```{r}
df <- read.csv(here("data/processed/reviews_with_sent_scores.csv"))
#view(df)
summary(df)
```
# Logistic Regression model:
# Create a subset for Logistic regression and Convert categorical variables:
```{r}
log_data <- df %>% 
  select( 
    flags_recommend_frend, 
    rating_culture_values, 
    rating_work_life, 
    rating_compensation_benefits, 
    rating_senior_leadership, 
    rating_career_opportunities, 
    rating_diversity_inclusion, 
    flags_business_outlook, 
    flags_ceo_approval,
    employee_length, 
    employee_job_quit, 
    sector, 
    covid_related_review, 
    career_missing
  )

# Recode recommendation / outlook / CEO approval into numeric variables
log_data <- log_data %>%
  dplyr::mutate(
    employee_job_quit = as.integer(employee_job_quit),
    career_missing = as.integer(career_missing),
    flags_recommend_frend = as.factor(flags_recommend_frend),
    sector = as.integer(as.factor(sector)),
    flags_business_outlook = as.integer(as.factor(flags_business_outlook)),
    flags_ceo_approval = as.integer(as.factor(flags_ceo_approval))
  )
```


# Split the data into training-testing
```{r}
library(caret)

set.seed(123)

train_index <- createDataPartition(
  log_data$flags_recommend_frend,
  p = 0.8,
  list = FALSE
)

train.data <- log_data[train_index, ]
test.data <- log_data[-train_index, ]

# Check class imbalance after splitting
prop.table(table(train.data$flags_recommend_frend))
prop.table(table(test.data$flags_recommend_frend))
```
```{r}
# Fit the model
set.seed(123)
model_logit <- glm(
  flags_recommend_frend ~.,
  data = train.data,
  family = binomial
)

summary(model_logit)
```

```{r}
# Predict on test data
log_pred_prob <- predict(model_logit, test.data, type = "response")
log_pred_prob
```

# Evaluate different classification thresholds to select the best one using F1 scores
# F1 = 2 * (precision * recall) / (precision + recall)
```{r}
# Step 5.1: Create a Sequence of Thresholds
thresholds <- seq(0.1, 1.0, by = 0.1)

# Step 5.2: Define a function to calculate F1 score
f1_score <- function(actual, predicted) {

  tp <- sum(actual == "POSITIVE" & predicted == "POSITIVE")
  fp <- sum(actual == "NEGATIVE" & predicted == "POSITIVE")
  fn <- sum(actual == "POSITIVE" & predicted == "NEGATIVE")

  precision <- ifelse((tp + fp) == 0, 0, tp / (tp + fp))
  recall <- ifelse((tp + fn) == 0, 0, tp / (tp + fn))

  if (precision + recall == 0) return(0)

  f1 <- 2 * precision * recall / (precision + recall)

  return(f1)
}

# Step 5.3: Loop Through Thresholds
f1_results <- data.frame(threshold = thresholds, F1 = NA)

for (i in seq_along(thresholds)) {
  
  t <- thresholds[i]
  
  pred_class <- ifelse(log_pred_prob > t, "POSITIVE", "NEGATIVE")
  
  f1_results$F1[i] <- f1_score(test.data$flags_recommend_frend, pred_class)
}

# Step 5.4: Print F1 scores
print(f1_results)

# Step 5.5: Choose the threshold with the best F1 score
best_threshold <- f1_results$threshold[which.max(f1_results$F1)]
best_f1 <- max(f1_results$F1)

best_threshold
best_f1
```
# Convert Probabilities to Class Predictions using threshold of 0.4
```{r}
log_pred_class <- ifelse(log_pred_prob > 0.4,"POSITIVE", "NEGATIVE")

# Convert to factor
log_pred_class <- as.factor(log_pred_class)
```

# Evaluate model performance
```{r}
logistic_cm <- confusionMatrix(log_pred_class, test.data$flags_recommend_frend)
logistic_cm
```

# ROC Curve and AUC
```{r}
library(pROC)

log_roc_curve <- roc(test.data$flags_recommend_frend, log_pred_prob)
plot(log_roc_curve)

auc(log_roc_curve)
```
## Function to evaluate each regularized model (Lasso, Ridge, Elastic Net)
```{r}
evaluate_model <- function(prob, actual, model_name, thresholds = seq(0.1, 1.0, by = 0.05)) {
  
  f1_results <- data.frame(threshold = thresholds, F1 = NA)
  
  for (i in seq_along(thresholds)) {
    t <- thresholds[i]
    pred_class <- ifelse(prob > t, "POSITIVE", "NEGATIVE")
    f1_results$F1[i] <- f1_score(actual, pred_class)
  }
  
  best_threshold <- f1_results$threshold[which.max(f1_results$F1)]
  cat("Best threshold for", model_name, ":", best_threshold, "\n")
  
  pred_class <- ifelse(prob > best_threshold, "POSITIVE", "NEGATIVE")
  pred_class <- as.factor(pred_class)
  
  cm <- confusionMatrix(pred_class, actual)
  print(cm)
  
  roc_obj <- roc(actual, prob)
  plot(roc_obj, main = paste("ROC Curve -", model_name))
  
  model_auc <- auc(roc_obj)
  cat(model_name, "AUC:", model_auc, "\n")
  
  list(
    f1_results = f1_results,
    best_threshold = best_threshold,
    pred_class = pred_class,
    confusion_matrix = cm,
    roc = roc_obj,
    auc = model_auc
  )
}
```


## Lasso
```{r}
x_train <- model.matrix(flags_recommend_frend ~ ., data = train.data)[, -1]
x_test  <- model.matrix(flags_recommend_frend ~ ., data = test.data)[, -1]

y_train <- ifelse(train.data$flags_recommend_frend == "POSITIVE", 1, 0)

cv_lasso <- cv.glmnet(
  x_train,
  y_train,
  family = "binomial",
  alpha = 1
)

lasso_prob <- predict(
  cv_lasso,
  newx = x_test,
  s = "lambda.min",
  type = "response"
)

lasso_prob <- as.numeric(lasso_prob)

lasso_eval <- evaluate_model(
  prob = lasso_prob,
  actual = test.data$flags_recommend_frend,
  model_name = "LASSO"
)
```

## Ridge
```{r}
set.seed(123)
cv_ridge <- cv.glmnet(
  x_train, y_train,
  family = "binomial",
  alpha = 0
)

ridge_prob <- predict(cv_ridge, newx = x_test, s = "lambda.min", type = "response")
ridge_prob <- as.numeric(ridge_prob)

ridge_eval <- evaluate_model(
  prob = ridge_prob,
  actual = test.data$flags_recommend_frend,
  model_name = "Ridge"
)
```

## Elastic Net
```{r}
set.seed(123)
cv_enet <- cv.glmnet(
  x_train, y_train,
  family = "binomial",
  alpha = 0.5
)

enet_prob <- predict(cv_enet, newx = x_test, s = "lambda.min", type = "response")
enet_prob <- as.numeric(enet_prob)

enet_eval <- evaluate_model(
  prob = enet_prob,
  actual = test.data$flags_recommend_frend,
  model_name = "Elastic Net"
)
```
## Logit again
```{r}
set.seed(123)

model_logit <- glm(
  flags_recommend_frend ~ .,
  data = train.data,
  family = binomial
)

logit_prob <- predict(
  model_logit,
  newdata = test.data,
  type = "response"
)

logit_eval <- evaluate_model(
  prob = log_pred_prob,
  actual = test.data$flags_recommend_frend,
  model_name = "Logistic"
)

```

```{r}
model_comparison <- data.frame(
  Model = c("Logistic", "LASSO", "Ridge", "Elastic Net"),
  Best_Threshold = c(
    logit_eval$best_threshold,
    lasso_eval$best_threshold,
    ridge_eval$best_threshold,
    enet_eval$best_threshold
  ),
  AUC = c(
    as.numeric(logit_eval$auc),
    as.numeric(lasso_eval$auc),
    as.numeric(ridge_eval$auc),
    as.numeric(enet_eval$auc)
  ),
  Accuracy = c(
    as.numeric(logit_eval$confusion_matrix$overall["Accuracy"]),
    as.numeric(lasso_eval$confusion_matrix$overall["Accuracy"]),
    as.numeric(ridge_eval$confusion_matrix$overall["Accuracy"]),
    as.numeric(enet_eval$confusion_matrix$overall["Accuracy"])
  ),
  F1 = c(
    max(logit_eval$f1_results$F1),
    max(lasso_eval$f1_results$F1),
    max(ridge_eval$f1_results$F1),
    max(enet_eval$f1_results$F1)
  )
)

model_comparison
```
## Sentiment Add-Ons

Calculate mean and sds across each model's text columns to get the average or compound sentiment
```{r}
df$sent_vader_avg <- rowMeans(
  df[, c(
    "vader_summary_compound",
    "vader_pros_compound",
    "vader_cons_compound",
    "vader_advice_compound"
  )],
  na.rm = TRUE
)

df$sent_vader_sd <- apply(
  df[, c(
    "vader_summary_compound",
    "vader_pros_compound",
    "vader_cons_compound",
    "vader_advice_compound"
  )],
  1,
  sd,
  na.rm = TRUE
)

df$sent_distilbert_avg <- rowMeans(
  df[, c(
    "distilbert_summary_signed",
    "distilbert_pros_signed",
    "distilbert_cons_signed",
    "distilbert_advice_signed"
  )],
  na.rm = TRUE
)

df$sent_distilbert_sd <- apply(
  df[, c(
    "distilbert_summary_signed",
    "distilbert_pros_signed",
    "distilbert_cons_signed",
    "distilbert_advice_signed"
  )],
  1,
  sd,
  na.rm = TRUE
)

df$sent_roberta_avg <- rowMeans(
  df[, c(
    "roberta_summary_signed",
    "roberta_pros_signed",
    "roberta_cons_signed",
    "roberta_advice_signed"
  )],
  na.rm = TRUE
)

df$sent_roberta_sd <- apply(
  df[, c(
    "roberta_summary_signed",
    "roberta_pros_signed",
    "roberta_cons_signed",
    "roberta_advice_signed"
  )],
  1,
  sd,
  na.rm = TRUE
)
```

# Just column binding the new compounded sentiment scores to original log_data
```{r}
log_data_vader <- cbind(
  log_data,
  df[, c(
    "sent_vader_avg",
    "sent_vader_sd"
  )]
)

log_data_distilbert <- cbind(
  log_data,
  df[, c(
    "sent_distilbert_avg",
    "sent_distilbert_sd"
  )]
)

log_data_roberta <- cbind(
  log_data,
  df[, c(
    "sent_roberta_avg",
    "sent_roberta_sd"
  )]
)

log_data_all <- cbind(
  log_data,
  df[, c(
    "sent_vader_avg",
    "sent_vader_sd",
    "sent_distilbert_avg",
    "sent_distilbert_sd",
    "sent_roberta_avg",
    "sent_roberta_sd"
  )]
)

datasets <- list(
  base_only = log_data,
  vader = log_data_vader,
  distilbert = log_data_distilbert,
  roberta = log_data_roberta,
  base_and_all_sent = log_data_all
)


```

```{r}
results <- list()

for (name in names(datasets)) {
  
  df_current <- datasets[[name]]
  
  train_df <- df_current[train_index, ]
  test_df  <- df_current[-train_index, ]
  
  # Logistic
  model_logit <- glm(
    flags_recommend_frend ~ .,
    data = train_df,
    family = binomial
  )
  
  logit_prob <- predict(model_logit, newdata = test_df, type = "response")
  
  logit_eval <- evaluate_model(
    prob = logit_prob,
    actual = test_df$flags_recommend_frend,
    model_name = paste("Base Logistic +", name)
  )
  
  # glmnet inputs
  x_train <- model.matrix(flags_recommend_frend ~ ., data = train_df)[, -1]
  x_test  <- model.matrix(flags_recommend_frend ~ ., data = test_df)[, -1]
  y_train <- ifelse(train_df$flags_recommend_frend == "POSITIVE", 1, 0)
  
  # LASSO
  cv_lasso <- cv.glmnet(x_train, y_train, family = "binomial", alpha = 1)
  lasso_prob <- as.numeric(predict(cv_lasso, newx = x_test, s = "lambda.min", type = "response"))
  
  lasso_eval <- evaluate_model(
    prob = lasso_prob,
    actual = test_df$flags_recommend_frend,
    model_name = paste("LASSO +", name)
  )
  
  # Ridge
  cv_ridge <- cv.glmnet(x_train, y_train, family = "binomial", alpha = 0)
  ridge_prob <- as.numeric(predict(cv_ridge, newx = x_test, s = "lambda.min", type = "response"))
  
  ridge_eval <- evaluate_model(
    prob = ridge_prob,
    actual = test_df$flags_recommend_frend,
    model_name = paste("Ridge +", name)
  )
  
  # Elastic Net
  cv_enet <- cv.glmnet(x_train, y_train, family = "binomial", alpha = 0.5)
  enet_prob <- as.numeric(predict(cv_enet, newx = x_test, s = "lambda.min", type = "response"))
  
  enet_eval <- evaluate_model(
    prob = enet_prob,
    actual = test_df$flags_recommend_frend,
    model_name = paste("Elastic Net +", name)
  )
  
  # Store results
  results[[name]] <- list(
    logit = logit_eval,
    lasso = lasso_eval,
    ridge = ridge_eval,
    enet  = enet_eval
  )
}


```

```{r}
model_comparison <- do.call(rbind, lapply(names(results), function(name) {
  
  res <- results[[name]]
  
  data.frame(
    Features = name,
    Model = c("Logistic", "LASSO", "Ridge", "Elastic Net"),
    
    AUC = c(
      as.numeric(res$logit$auc),
      as.numeric(res$lasso$auc),
      as.numeric(res$ridge$auc),
      as.numeric(res$enet$auc)
    ),
    
    F1 = c(
      max(res$logit$f1_results$F1),
      max(res$lasso$f1_results$F1),
      max(res$ridge$f1_results$F1),
      max(res$enet$f1_results$F1)
    ),
    
    Best_Threshold = c(
      res$logit$best_threshold,
      res$lasso$best_threshold,
      res$ridge$best_threshold,
      res$enet$best_threshold
    ),
    
    Accuracy = c(
      as.numeric(res$logit$confusion_matrix$overall["Accuracy"]),
      as.numeric(res$lasso$confusion_matrix$overall["Accuracy"]),
      as.numeric(res$ridge$confusion_matrix$overall["Accuracy"]),
      as.numeric(res$enet$confusion_matrix$overall["Accuracy"])
    )
  )
}))

model_comparison %>% arrange(desc(AUC))
```


# Decision Tree

# Import packages
```{r}
# ---- 0. Install / load packages ----
pkg_needed <- c("caret", "rpart", "rattle")
for (p in pkg_needed) {
  if (!requireNamespace(p, quietly = TRUE)) {
    install.packages(p)
  }
}

#install.packages("bitops")
#install.packages("ggplot2")
#install.packages("lattice")
#install.packages("Matrix")
#install.packages("rpart.plot")
library(caret)
library(rpart)
library(rattle)
library(rpart.plot)
```

# Create a subset for decision tree 
```{r}
tree_data <- df %>% 
  select( 
    flags_recommend_frend, 
    rating_overall,
    rating_culture_values, 
    rating_work_life, 
    rating_compensation_benefits, 
    rating_senior_leadership, 
    rating_career_opportunities, 
    rating_diversity_inclusion, 
    flags_business_outlook, 
    flags_ceo_approval,
    employee_length, 
    employee_job_quit, 
    sector, 
    covid_related_review, 
    career_missing
  )

# Recode recommendation / outlook / CEO approval into numeric variables
tree_data <- tree_data %>%
  dplyr::mutate(
    employee_job_quit = as.integer(employee_job_quit),
    career_missing = as.integer(career_missing),
    flags_recommend_frend = as.factor(flags_recommend_frend),
    sector = as.integer(as.factor(sector)),
    flags_business_outlook = as.integer(as.factor(flags_business_outlook)),
    flags_ceo_approval = as.integer(as.factor(flags_ceo_approval))
  )
```

# Split the dataset into training and testing
```{r}
library(caret)

set.seed(123)

train_index <- createDataPartition(
  tree_data$flags_recommend_frend,
  p = 0.8,
  list = FALSE
)

tree.train.data <- tree_data[train_index, ]
tree.test.data <- tree_data[-train_index, ]

# Check class imbalance after splitting
prop.table(table(tree.train.data$flags_recommend_frend))
prop.table(table(tree.test.data$flags_recommend_frend))
```

# Train the Decision Tree
```{r}
# Fit the initial tree
set.seed(123)
tree_model <- rpart(
  flags_recommend_frend ~ .,
  data = tree.train.data,
  method = "class", 
  control = rpart.control(cp = 0, xval = 10, minsplit = 2, minbucket = 1)
)

# Visualize the tree
rpart.plot(tree_model, main = "Initial Tree")

# Check CP table
printcp(tree_model)

# Select the best CP value
best_cp <- tree_model$cptable[
  which.min(tree_model$cptable[, "xerror"]),
  "CP"
]

cat("\nBest CP value:", best_cp, "\n")

# Prune the tree using best CP value
pruned_tree <- prune(tree_model, cp = best_cp)

# Visualize the pruned tree
fancyRpartPlot(pruned_tree, main = "Optimal Pruned Tree")

# Predict on test data
pred_test <- predict(pruned_tree, newdata = tree.test.data, type = "class")

# Confusion Matrix and Metrics
pruned_tree_cm <- confusionMatrix(pred_test, tree.test.data$flags_recommend_frend)
pruned_tree_cm
```
# ROC curve and AUC 
```{r}
# ROC and AUC for pruned tree
#install.packages("pROC")
library(pROC)

# Make sure outcome levels are ordered correctly
test.data$flags_recommend_frend <- factor(
  tree.test.data$flags_recommend_frend,
  levels = c("NEGATIVE", "POSITIVE")
)

# Predicted probabilities
pred_prob <- predict(pruned_tree, newdata = tree.test.data, type = "prob")

# Probability of positive class
pred_prob_positive <- pred_prob[, "POSITIVE"]

# ROC curve
roc_pruned_tree <- roc(
  response = tree.test.data$flags_recommend_frend,
  predictor = pred_prob_positive,
  levels = c("NEGATIVE", "POSITIVE"),
  direction = "<"
)

# AUC
auc_pruned_tree <- auc(roc_pruned_tree)
auc_pruned_tree

# Plot ROC
plot(
  roc_pruned_tree,
  col = "blue",
  lwd = 2,
  main = "ROC Curve - Pruned Decision Tree"
)

# Add diagonal reference line
abline(a = 0, b = 1, lty = 2, col = "gray")
```
# Random forest:

# Import packages
```{r}
#install.packages("randomForest")
#install.packages("ggplot2")
library(randomForest)
library(caret)
```


# Train the random forest
```{r}
# Train Random Forest with 10-fold CV
set.seed(123)

p <- ncol(tree.train.data) - 1

rf_model <- train(
  flags_recommend_frend ~ .,
  data = tree.train.data,
  method = "rf",
  trControl = trainControl(method = "cv", number = 10),
  tuneGrid = expand.grid(mtry = 1:p),
  ntree = 500,
  importance = TRUE
)

# Find the best mtry
bestTune <- rf_model$bestTune
bestTune

# Refit the model with the best mtry
rf_model_tuned <- train(
  flags_recommend_frend ~ .,
  data = tree.train.data,
  method = "rf",
  trControl = trainControl(
    method = "cv",
    number = 10,
    classProbs = TRUE,
    summaryFunction = twoClassSummary
  ),
  tuneGrid = expand.grid(mtry = bestTune),  # use your chosen mtry
  ntree = 500,
  metric = "ROC",
  importance = TRUE
)

# Variable importance plots
varImpPlot(rf_model_tuned$finalModel, type = 1, main = "Mean Decrease Accuracy (MSE)")  # Mean Decrease Accuracy (MSE)
varImpPlot(rf_model_tuned$finalModel, type = 2, main = "Mean Decrease Node Impurity")  # Mean Decrease Node Impurity

# Predict on test data
pred_test <- predict(rf_model_tuned, tree.test.data)

# Confusion matrix and metrics
cm <- confusionMatrix(
  pred_test,
  tree.test.data$flags_recommend_frend
)

cm

# Metrics
sensitivity <- cm$byClass["Sensitivity"]
specificity <- cm$byClass["Specificity"]
accuracy    <- cm$overall["Accuracy"]

sensitivity
specificity
accuracy
```
# ROC curve and AUC
```{r}
library(pROC)
# Predict probabilities
rf_pred_prob <- predict(rf_model_tuned, newdata = tree.test.data, type = "prob")

# Extract probability of POSITIVE class
rf_pred_prob_pos <- rf_pred_prob[, "POSITIVE"]

# Make Sure Target is Factor
test.data$flags_recommend_frend <- factor(
  tree.test.data$flags_recommend_frend,
  levels = c("NEGATIVE", "POSITIVE")
)

# Compute ROC Curve
roc_rf <- roc(
  response = tree.test.data$flags_recommend_frend,
  predictor = rf_pred_prob_pos,
  levels = c("NEGATIVE", "POSITIVE"),
  direction = "<"
)

# AUC
auc_rf <- auc(roc_rf)
auc_rf

# Plot ROC curve
plot(
  roc_rf,
  col = "blue",
  lwd = 2,
  main = "ROC Curve - Random Forest"
)

abline(a = 0, b = 1, lty = 2, col = "gray")
```

