---
title: "modeling_with_cleaned_data"
author: "Quynh Tran"
date: "2026-04-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import packages
```{r}
library(MASS)
library(tidyverse)
#install.packages("lattice")
library(lattice)
library(caret)
#install.packages("Matrix")
library(Matrix)
library(glmnet)
library(mlbench)
library(caret)
library(here)
```

# Import dataset
```{r}
df <- read.csv(here("data/processed/reviews_with_sent_scores.csv"))
#view(df)
summary(df)
```
# Logistic Regression model:
# Create a subset for Logistic regression and Convert categorical variables:
```{r}
log_data <- df %>% 
  select( 
    flags_recommend_frend, 
    rating_culture_values, 
    rating_work_life, 
    rating_compensation_benefits, 
    rating_senior_leadership, 
    rating_career_opportunities, 
    rating_diversity_inclusion, 
    flags_business_outlook, 
    flags_ceo_approval,
    employee_length, 
    employee_job_quit, 
    sector, 
    covid_related_review, 
    career_missing
  )

# Recode recommendation / outlook / CEO approval into numeric variables
log_data <- log_data %>%
  dplyr::mutate(
    employee_job_quit = as.integer(employee_job_quit),
    career_missing = as.integer(career_missing),
    flags_recommend_frend = as.factor(flags_recommend_frend),
    sector = as.integer(as.factor(sector)),
    flags_business_outlook = as.integer(as.factor(flags_business_outlook)),
    flags_ceo_approval = as.integer(as.factor(flags_ceo_approval))
  )
```


# Split the data into training-testing
```{r}
library(caret)

set.seed(123)

train_index <- createDataPartition(
  log_data$flags_recommend_frend,
  p = 0.8,
  list = FALSE
)

train.data <- log_data[train_index, ]
test.data <- log_data[-train_index, ]

# Check class imbalance after splitting
prop.table(table(train.data$flags_recommend_frend))
prop.table(table(test.data$flags_recommend_frend))
```
```{r}
# Fit the model
set.seed(123)
model_logit <- glm(
  flags_recommend_frend ~.,
  data = train.data,
  family = binomial
)

summary(model_logit)
```

```{r}
# Predict on test data
log_pred_prob <- predict(model_logit, test.data, type = "response")
log_pred_prob
```

# Evaluate different classification thresholds to select the best one using F1 scores
# F1 = 2 * (precision * recall) / (precision + recall)
```{r}
# Step 5.1: Create a Sequence of Thresholds
thresholds <- seq(0.1, 1.0, by = 0.1)

# Step 5.2: Define a function to calculate F1 score
f1_score <- function(actual, predicted) {

  tp <- sum(actual == "POSITIVE" & predicted == "POSITIVE")
  fp <- sum(actual == "NEGATIVE" & predicted == "POSITIVE")
  fn <- sum(actual == "POSITIVE" & predicted == "NEGATIVE")

  precision <- ifelse((tp + fp) == 0, 0, tp / (tp + fp))
  recall <- ifelse((tp + fn) == 0, 0, tp / (tp + fn))

  if (precision + recall == 0) return(0)

  f1 <- 2 * precision * recall / (precision + recall)

  return(f1)
}

# Step 5.3: Loop Through Thresholds
f1_results <- data.frame(threshold = thresholds, F1 = NA)

for (i in seq_along(thresholds)) {
  
  t <- thresholds[i]
  
  pred_class <- ifelse(log_pred_prob > t, "POSITIVE", "NEGATIVE")
  
  f1_results$F1[i] <- f1_score(test.data$flags_recommend_frend, pred_class)
}

# Step 5.4: Print F1 scores
print(f1_results)

# Step 5.5: Choose the threshold with the best F1 score
best_threshold <- f1_results$threshold[which.max(f1_results$F1)]
best_f1 <- max(f1_results$F1)

best_threshold
best_f1
```
# Convert Probabilities to Class Predictions using threshold of 0.4
```{r}
log_pred_class <- ifelse(log_pred_prob > 0.4,"POSITIVE", "NEGATIVE")

# Convert to factor
log_pred_class <- as.factor(log_pred_class)
```

# Evaluate model performance
```{r}
logistic_cm <- confusionMatrix(log_pred_class, test.data$flags_recommend_frend)
logistic_cm
```

# ROC Curve and AUC
```{r}
library(pROC)

log_roc_curve <- roc(test.data$flags_recommend_frend, log_pred_prob)
plot(log_roc_curve)

auc(log_roc_curve)
```
# Decision Tree

# Import packages
```{r}
# ---- 0. Install / load packages ----
pkg_needed <- c("caret", "rpart", "rattle")
for (p in pkg_needed) {
  if (!requireNamespace(p, quietly = TRUE)) {
    install.packages(p)
  }
}

#install.packages("bitops")
#install.packages("ggplot2")
#install.packages("lattice")
#install.packages("Matrix")
#install.packages("rpart.plot")
library(caret)
library(rpart)
library(rattle)
library(rpart.plot)
```

# Create a subset for decision tree 
```{r}
tree_data <- df %>% 
  select( 
    flags_recommend_frend, 
    rating_overall,
    rating_culture_values, 
    rating_work_life, 
    rating_compensation_benefits, 
    rating_senior_leadership, 
    rating_career_opportunities, 
    rating_diversity_inclusion, 
    flags_business_outlook, 
    flags_ceo_approval,
    employee_length, 
    employee_job_quit, 
    sector, 
    covid_related_review, 
    career_missing
  )

# Recode recommendation / outlook / CEO approval into numeric variables
tree_data <- tree_data %>%
  dplyr::mutate(
    employee_job_quit = as.integer(employee_job_quit),
    career_missing = as.integer(career_missing),
    flags_recommend_frend = as.factor(flags_recommend_frend),
    sector = as.integer(as.factor(sector)),
    flags_business_outlook = as.integer(as.factor(flags_business_outlook)),
    flags_ceo_approval = as.integer(as.factor(flags_ceo_approval))
  )
```

# Split the dataset into training and testing
```{r}
library(caret)

set.seed(123)

train_index <- createDataPartition(
  tree_data$flags_recommend_frend,
  p = 0.8,
  list = FALSE
)

tree.train.data <- tree_data[train_index, ]
tree.test.data <- tree_data[-train_index, ]

# Check class imbalance after splitting
prop.table(table(tree.train.data$flags_recommend_frend))
prop.table(table(tree.test.data$flags_recommend_frend))
```

# Train the Decision Tree
```{r}
# Fit the initial tree
set.seed(123)
tree_model <- rpart(
  flags_recommend_frend ~ .,
  data = tree.train.data,
  method = "class", 
  control = rpart.control(cp = 0, xval = 10, minsplit = 2, minbucket = 1)
)

# Visualize the tree
rpart.plot(tree_model, main = "Initial Tree")

# Check CP table
printcp(tree_model)

# Select the best CP value
best_cp <- tree_model$cptable[
  which.min(tree_model$cptable[, "xerror"]),
  "CP"
]

cat("\nBest CP value:", best_cp, "\n")

# Prune the tree using best CP value
pruned_tree <- prune(tree_model, cp = best_cp)

# Visualize the pruned tree
fancyRpartPlot(pruned_tree, main = "Optimal Pruned Tree")

# Predict on test data
pred_test <- predict(pruned_tree, newdata = tree.test.data, type = "class")

# Confusion Matrix and Metrics
pruned_tree_cm <- confusionMatrix(pred_test, tree.test.data$flags_recommend_frend)
pruned_tree_cm
```
# ROC curve and AUC 
```{r}
# ROC and AUC for pruned tree
#install.packages("pROC")
library(pROC)

# Make sure outcome levels are ordered correctly
test.data$flags_recommend_frend <- factor(
  tree.test.data$flags_recommend_frend,
  levels = c("NEGATIVE", "POSITIVE")
)

# Predicted probabilities
pred_prob <- predict(pruned_tree, newdata = tree.test.data, type = "prob")

# Probability of positive class
pred_prob_positive <- pred_prob[, "POSITIVE"]

# ROC curve
roc_pruned_tree <- roc(
  response = tree.test.data$flags_recommend_frend,
  predictor = pred_prob_positive,
  levels = c("NEGATIVE", "POSITIVE"),
  direction = "<"
)

# AUC
auc_pruned_tree <- auc(roc_pruned_tree)
auc_pruned_tree

# Plot ROC
plot(
  roc_pruned_tree,
  col = "blue",
  lwd = 2,
  main = "ROC Curve - Pruned Decision Tree"
)

# Add diagonal reference line
abline(a = 0, b = 1, lty = 2, col = "gray")
```
# Random forest:

# Import packages
```{r}
#install.packages("randomForest")
#install.packages("ggplot2")
library(randomForest)
library(caret)
```


# Train the random forest
```{r}
# Train Random Forest with 10-fold CV
set.seed(123)

p <- ncol(tree.train.data) - 1

rf_model <- train(
  flags_recommend_frend ~ .,
  data = tree.train.data,
  method = "rf",
  trControl = trainControl(method = "cv", number = 10),
  tuneGrid = expand.grid(mtry = 1:p),
  ntree = 500,
  importance = TRUE
)

# Find the best mtry
bestTune <- rf_model$bestTune
bestTune

# Refit the model with the best mtry
rf_model_tuned <- train(
  flags_recommend_frend ~ .,
  data = tree.train.data,
  method = "rf",
  trControl = trainControl(
    method = "cv",
    number = 10,
    classProbs = TRUE,
    summaryFunction = twoClassSummary
  ),
  tuneGrid = expand.grid(mtry = bestTune),  # 👈 use your chosen mtry
  ntree = 500,
  metric = "ROC",
  importance = TRUE
)

# Variable importance plots
varImpPlot(rf_model_tuned$finalModel, type = 1, main = "Mean Decrease Accuracy (MSE)")  # Mean Decrease Accuracy (MSE)
varImpPlot(rf_model_tuned$finalModel, type = 2, main = "Mean Decrease Node Impurity")  # Mean Decrease Node Impurity

# Predict on test data
pred_test <- predict(rf_model_tuned, tree.test.data)

# Confusion matrix and metrics
cm <- confusionMatrix(
  pred_test,
  tree.test.data$flags_recommend_frend
)

cm

# Metrics
sensitivity <- cm$byClass["Sensitivity"]
specificity <- cm$byClass["Specificity"]
accuracy    <- cm$overall["Accuracy"]

sensitivity
specificity
accuracy
```
# ROC curve and AUC
```{r}
library(pROC)
# Predict probabilities
rf_pred_prob <- predict(rf_model_tuned, newdata = tree.test.data, type = "prob")

# Extract probability of POSITIVE class
rf_pred_prob_pos <- rf_pred_prob[, "POSITIVE"]

# Make Sure Target is Factor
test.data$flags_recommend_frend <- factor(
  tree.test.data$flags_recommend_frend,
  levels = c("NEGATIVE", "POSITIVE")
)

# Compute ROC Curve
roc_rf <- roc(
  response = tree.test.data$flags_recommend_frend,
  predictor = rf_pred_prob_pos,
  levels = c("NEGATIVE", "POSITIVE"),
  direction = "<"
)

# AUC
auc_rf <- auc(roc_rf)
auc_rf

# Plot ROC curve
plot(
  roc_rf,
  col = "blue",
  lwd = 2,
  main = "ROC Curve - Random Forest"
)

abline(a = 0, b = 1, lty = 2, col = "gray")
```

