---
title: "recommend_friend_models"
date: "2026-04-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(MASS)
library(tidyverse)
library(caret)
library(Matrix)
library(glmnet)
library(here)
library(pROC)
library(randomForest)
library(broom)
library(tidytext)

set.seed(123)

df <- read.csv(here("data/processed/reviews_with_sent_scores.csv"))
```

## Create Aggregated Sentiment Features
```{r}
df <- df %>%
  mutate(
    sent_vader_avg = rowMeans(
      across(c(
        vader_summary_compound,
        vader_pros_compound,
        vader_cons_compound,
        vader_advice_compound
      )),
      na.rm = TRUE
    ),
    
    sent_vader_sd = apply(
      select(., c(
        vader_summary_compound,
        vader_pros_compound,
        vader_cons_compound,
        vader_advice_compound
      )),
      1,
      sd,
      na.rm = TRUE
    ),
    
    sent_distilbert_avg = rowMeans(
      across(c(
        distilbert_summary_signed,
        distilbert_pros_signed,
        distilbert_cons_signed,
        distilbert_advice_signed
      )),
      na.rm = TRUE
    ),
    
    sent_distilbert_sd = apply(
      select(., c(
        distilbert_summary_signed,
        distilbert_pros_signed,
        distilbert_cons_signed,
        distilbert_advice_signed
      )),
      1,
      sd,
      na.rm = TRUE
    ),
    
    sent_roberta_avg = rowMeans(
      across(c(
        roberta_summary_signed,
        roberta_pros_signed,
        roberta_cons_signed,
        roberta_advice_signed
      )),
      na.rm = TRUE
    ),
    
    sent_roberta_sd = apply(
      select(., c(
        roberta_summary_signed,
        roberta_pros_signed,
        roberta_cons_signed,
        roberta_advice_signed
      )),
      1,
      sd,
      na.rm = TRUE
    )
  )
```

## Define Base Features and Helper Functions
```{r}
base_features <- c(
  "rating_overall",
  "rating_culture_values",
  "rating_work_life",
  "rating_compensation_benefits",
  "rating_senior_leadership",
  "rating_career_opportunities",
  "rating_diversity_inclusion",
  "flags_business_outlook",
  "flags_ceo_approval",
  "employee_length",
  "employee_job_quit",
  "sector",
  "covid_related_review",
  "career_missing"
  
)

make_model_data <- function(df, extra_cols = NULL) {
  df %>%
    dplyr::select(
      flags_recommend_frend,
      all_of(base_features),
      all_of(extra_cols)
    ) %>%
    mutate(
      employee_job_quit = as.integer(employee_job_quit),
      career_missing = as.integer(career_missing),
      flags_recommend_frend = factor(
        flags_recommend_frend,
        levels = c("NEGATIVE", "POSITIVE")
      ),
      sector = as.factor(sector),
      #flags_business_outlook = as.integer(as.factor(flags_business_outlook)),
      #flags_ceo_approval = as.integer(as.factor(flags_ceo_approval))
      flags_business_outlook = as.numeric(recode(flags_business_outlook,
                                "NEGATIVE" = -1,"NEUTRAL"  = 0, "POSITIVE" = 1)),
      flags_ceo_approval = as.factor(flags_ceo_approval)
    )
}

f1_score <- function(actual, predicted) {
  tp <- sum(actual == "POSITIVE" & predicted == "POSITIVE")
  fp <- sum(actual == "NEGATIVE" & predicted == "POSITIVE")
  fn <- sum(actual == "POSITIVE" & predicted == "NEGATIVE")
  
  precision <- ifelse((tp + fp) == 0, 0, tp / (tp + fp))
  recall <- ifelse((tp + fn) == 0, 0, tp / (tp + fn))
  
  if (precision + recall == 0) return(0)
  
  2 * precision * recall / (precision + recall)
}

evaluate_model <- function(prob, actual, thresholds = seq(0.1, 1.0, by = 0.05)) {
  f1_results <- tibble(threshold = thresholds) %>%
    mutate(
      F1 = purrr::map_dbl(threshold, ~{
        pred_class <- ifelse(prob > .x, "POSITIVE", "NEGATIVE")
        f1_score(actual, pred_class)
      })
    )
  
  best_threshold <- f1_results %>%
    slice_max(F1, n = 1, with_ties = FALSE) %>%
    pull(threshold)
  
  pred_class <- factor(
    ifelse(prob > best_threshold, "POSITIVE", "NEGATIVE"),
    levels = c("NEGATIVE", "POSITIVE")
  )
  
  cm <- confusionMatrix(pred_class, actual)
  roc_obj <- roc(actual, prob, levels = c("NEGATIVE", "POSITIVE"), direction = "<")
  
  list(
    f1_results = f1_results,
    best_threshold = best_threshold,
    pred_class = pred_class,
    confusion_matrix = cm,
    roc = roc_obj,
    auc = as.numeric(auc(roc_obj))
  )
}

extract_glm_importance <- function(model, features_name, model_name, top_n = 10) {
  broom::tidy(model) %>%
    dplyr::filter(term != "(Intercept)") %>%
    dplyr::mutate(
      Features = features_name,
      Model = model_name,
      Importance = abs(estimate),
      Direction = ifelse(estimate >= 0, "Positive", "Negative")
    ) %>%
    dplyr::arrange(desc(Importance)) %>%
    dplyr::slice_head(n = top_n) %>%
    dplyr::select(Features, Model, term, estimate, Importance, Direction)
}

extract_glmnet_importance <- function(cv_model, features_name, model_name, s = "lambda.min", top_n = 10) {
  coef_df <- as.matrix(coef(cv_model, s = s)) %>%
    as.data.frame() %>%
    tibble::rownames_to_column("term")
  
  names(coef_df)[2] <- "estimate"
  
  coef_df %>%
    dplyr::filter(term != "(Intercept)", estimate != 0) %>%
    dplyr::mutate(
      Features = features_name,
      Model = model_name,
      Importance = abs(estimate),
      Direction = ifelse(estimate >= 0, "Positive", "Negative")
    ) %>%
    dplyr::arrange(desc(Importance)) %>%
    dplyr::slice_head(n = top_n) %>%
    dplyr::select(Features, Model, term, estimate, Importance, Direction)
}

extract_rf_importance <- function(rf_model_tuned, features_name, model_name, top_n = 10) {
  imp <- importance(rf_model_tuned$finalModel, type = 1)
  
  data.frame(
    term = rownames(imp),
    Importance = imp[, 1],
    stringsAsFactors = FALSE
  ) %>%
    dplyr::mutate(
      Features = features_name,
      Model = model_name,
      estimate = Importance,
      Direction = "RF Importance"
    ) %>%
    dplyr::arrange(desc(Importance)) %>%
    dplyr::slice_head(n = top_n) %>%
    dplyr::select(Features, Model, term, estimate, Importance, Direction)
}

clean_feature_names <- function(x) {
  x %>%
    stringr::str_replace_all("_", " ") %>%
    stringr::str_replace_all("rating career opportunities", "career opportunities") %>%
    stringr::str_replace_all("rating compensation benefits", "comp benefits") %>%
    stringr::str_replace_all("rating senior leadership", "senior leadership") %>%
    stringr::str_replace_all("rating culture values", "culture values") %>%
    stringr::str_replace_all("rating work life", "work life") %>%
    stringr::str_replace_all("rating diversity inclusion", "diversity inclusion") %>%
    stringr::str_replace_all("flags business outlook", "business outlook") %>%
    stringr::str_replace_all("flags ceo approval", "ceo approval") %>%
    stringr::str_replace_all("employee job quit", "job quit") %>%
    stringr::str_replace_all("sent vader avg", "vader avg") %>%
    stringr::str_replace_all("sent vader sd", "vader sd") %>%
    stringr::str_replace_all("sent distilbert avg", "distilbert avg") %>%
    stringr::str_replace_all("sent distilbert sd", "distilbert sd") %>%
    stringr::str_replace_all("sent roberta avg", "roberta avg") %>%
    stringr::str_replace_all("sent roberta sd", "roberta sd")
}
```




## Build Datasets
```{r}
datasets <- list(
  base_only = make_model_data(df),
  vader = make_model_data(df, c("sent_vader_avg", "sent_vader_sd")),
  distilbert = make_model_data(df, c("sent_distilbert_avg", "sent_distilbert_sd")),
  roberta = make_model_data(df, c("sent_roberta_avg", "sent_roberta_sd")),
  base_and_all_sent = make_model_data(df, c(
    "sent_vader_avg", "sent_vader_sd",
    "sent_distilbert_avg", "sent_distilbert_sd",
    "sent_roberta_avg", "sent_roberta_sd"
  ))
)

train_index <- createDataPartition(
  datasets$base_only$flags_recommend_frend,
  p = 0.8,
  list = FALSE
)
```

# 1. Fit Base-Only Logistic Models
```{r}
base_train <- datasets$base_only[train_index, ]
base_test  <- datasets$base_only[-train_index, ]

x_train_base <- model.matrix(flags_recommend_frend ~ ., data = base_train)[, -1]
x_test_base  <- model.matrix(flags_recommend_frend ~ ., data = base_test)[, -1]
y_train_base <- ifelse(base_train$flags_recommend_frend == "POSITIVE", 1, 0)

base_logit <- glm(
  flags_recommend_frend ~ .,
  data = base_train,
  family = binomial
)
base_logit_prob <- predict(base_logit, newdata = base_test, type = "response")
base_logit_eval <- evaluate_model(base_logit_prob, base_test$flags_recommend_frend)

base_lasso <- cv.glmnet(
  x_train_base,
  y_train_base,
  family = "binomial",
  alpha = 1
)
base_lasso_prob <- as.numeric(
  predict(base_lasso, newx = x_test_base, s = "lambda.min", type = "response")
)
base_lasso_eval <- evaluate_model(base_lasso_prob, base_test$flags_recommend_frend)

base_ridge <- cv.glmnet(
  x_train_base,
  y_train_base,
  family = "binomial",
  alpha = 0
)
base_ridge_prob <- as.numeric(
  predict(base_ridge, newx = x_test_base, s = "lambda.min", type = "response")
)
base_ridge_eval <- evaluate_model(base_ridge_prob, base_test$flags_recommend_frend)

base_enet <- cv.glmnet(
  x_train_base,
  y_train_base,
  family = "binomial",
  alpha = 0.5
)
base_enet_prob <- as.numeric(
  predict(base_enet, newx = x_test_base, s = "lambda.min", type = "response")
)
base_enet_eval <- evaluate_model(base_enet_prob, base_test$flags_recommend_frend)
```

# 1.1 Summarize Base-Only Logistic Models
```{r}
base_logit_summary <- tibble(
  Features = "base_only",
  Model = c("Logistic Regression", "LASSO LogReg", "Ridge LogReg", "Elastic Net LogReg"),
  AUC = c(
    base_logit_eval$auc,
    base_lasso_eval$auc,
    base_ridge_eval$auc,
    base_enet_eval$auc
  ),
  F1 = c(
    max(base_logit_eval$f1_results$F1),
    max(base_lasso_eval$f1_results$F1),
    max(base_ridge_eval$f1_results$F1),
    max(base_enet_eval$f1_results$F1)
  ),
  Accuracy = c(
    as.numeric(base_logit_eval$confusion_matrix$overall["Accuracy"]),
    as.numeric(base_lasso_eval$confusion_matrix$overall["Accuracy"]),
    as.numeric(base_ridge_eval$confusion_matrix$overall["Accuracy"]),
    as.numeric(base_enet_eval$confusion_matrix$overall["Accuracy"])
  ),
  Best_Threshold = c(
    base_logit_eval$best_threshold,
    base_lasso_eval$best_threshold,
    base_ridge_eval$best_threshold,
    base_enet_eval$best_threshold
  )
)

base_logit_summary
```

# 1.3 Plot Base-Only Logistic Models by AUC Rank, F1 Rank, and Average Rank

```{r plot-base-only-logistic-ranks, fig.width=9, fig.height=5}
base_logit_rank_plot <- base_logit_summary %>%
  mutate(
    Model_Label = Model,
    AUC_Rank = dense_rank(desc(AUC)),
    F1_Rank = dense_rank(desc(F1)),
    Avg_Rank = (AUC_Rank + F1_Rank) / 2
  ) %>%
  arrange(Avg_Rank) %>%
  mutate(Model_Label = factor(Model_Label, levels = rev(Model_Label)))

ggplot(base_logit_rank_plot, aes(y = Model_Label)) +
  geom_segment(
    aes(x = AUC_Rank, xend = F1_Rank, yend = Model_Label),
    color = "gray70",
    linewidth = 1
  ) +
  geom_point(aes(x = AUC_Rank, color = "AUC Rank"), size = 3) +
  geom_point(aes(x = F1_Rank, color = "F1 Rank"), size = 3) +
  geom_point(aes(x = Avg_Rank, color = "Average Rank"), size = 3.5, shape = 18) +
  scale_x_reverse() +
  theme_minimal() +
  labs(
    title = "Base-Only Logistic Models: AUC Rank vs F1 Rank vs Average Rank",
    x = "Rank (1 = Best)",
    y = NULL,
    color = "Metric"
  )
```
# 2. Fit Base-Only Random Forest
```{r}
p_base <- ncol(base_train) - 1

rf_base_tune <- train(
  flags_recommend_frend ~ .,
  data = base_train,
  method = "rf",
  trControl = trainControl(method = "cv", number = 10),
  tuneGrid = expand.grid(mtry = 1:p_base),
  ntree = 500,
  importance = TRUE
)

rf_base_best <- rf_base_tune$bestTune

rf_base <- train(
  flags_recommend_frend ~ .,
  data = base_train,
  method = "rf",
  trControl = trainControl(
    method = "cv",
    number = 10,
    classProbs = TRUE,
    summaryFunction = twoClassSummary
  ),
  tuneGrid = rf_base_best,
  ntree = 500,
  metric = "ROC",
  importance = TRUE
)

rf_base_prob <- predict(rf_base, newdata = base_test, type = "prob")[, "POSITIVE"]
rf_base_eval <- evaluate_model(rf_base_prob, base_test$flags_recommend_frend)


rf_base_summary <- tibble(
  Features = "base_only",
  Model = "Random Forest",
  AUC = rf_base_eval$auc,
  F1 = max(rf_base_eval$f1_results$F1),
  Accuracy = as.numeric(rf_base_eval$confusion_matrix$overall["Accuracy"]),
  Best_Threshold = rf_base_eval$best_threshold
)

rf_base_summary
```

# 3 Compare Base-Only Models
```{r}
base_only_summary <- bind_rows(
  base_logit_summary %>% dplyr::select(Features, Model, AUC, F1, Accuracy),
  rf_base_summary %>% dplyr::select(Features, Model, AUC, F1, Accuracy)
) %>%
  mutate(
    Model_Label = paste(Features, Model, sep = " + "),
    AUC_Rank = dense_rank(desc(AUC)),
    F1_Rank = dense_rank(desc(F1)),
    Avg_Rank = (AUC_Rank + F1_Rank) / 2
  ) %>%
  arrange(Avg_Rank)

base_only_summary
```

# 3.1 Plot Base-Only Model Comparison
```{r}
base_only_summary_plot <- base_only_summary %>%
  arrange(Avg_Rank) %>%
  mutate(Model_Label = factor(Model_Label, levels = rev(Model_Label)))

ggplot(base_only_summary_plot, aes(y = Model_Label)) +
  geom_segment(
    aes(x = AUC_Rank, xend = F1_Rank, yend = Model_Label),
    color = "gray70",
    linewidth = 1
  ) +
  geom_point(aes(x = AUC_Rank, color = "AUC Rank"), size = 3) +
  geom_point(aes(x = F1_Rank, color = "F1 Rank"), size = 3) +
  geom_point(aes(x = Avg_Rank, color = "Average Rank"), size = 3.5, shape = 18) +
  scale_x_reverse() +
  theme_minimal() +
  labs(
    title = "Base-Only Models: AUC Rank vs F1 Rank vs Average Rank",
    x = "Rank (1 = Best)",
    y = NULL,
    color = "Metric"
  )
```


# 4 Extract Top 10 Base-Only Features for Each Model
```{r}
base_logit_importance <- extract_glm_importance(
  base_logit,
  "base_only",
  "Logistic Regression",
  top_n = 10
)

base_lasso_importance <- extract_glmnet_importance(
  base_lasso,
  "base_only",
  "LASSO LogReg",
  top_n = 10
)

base_ridge_importance <- extract_glmnet_importance(
  base_ridge,
  "base_only",
  "Ridge LogReg",
  top_n = 10
)

base_enet_importance <- extract_glmnet_importance(
  base_enet,
  "base_only",
  "Elastic Net LogReg",
  top_n = 10
)

base_rf_importance <- extract_rf_importance(
  rf_base,
  "base_only",
  "Random Forest",
  top_n = 10
)

base_importance_all <- bind_rows(
  base_logit_importance,
  base_lasso_importance,
  base_ridge_importance,
  base_enet_importance,
  base_rf_importance
) %>%
  mutate(
    Model_Label = paste(Features, Model, sep = " + "),
    term_clean = clean_feature_names(term)
  )

base_importance_all
```

# 4.1 Plot Top 10 Features for Base-Only Logistic-Family Models

```{r plot-base-only-logit-top-features, fig.width=10, fig.height=8}
base_logit_features_df <- base_importance_all %>%
  filter(Model != "Random Forest") %>%
  group_by(Model_Label) %>%
  slice_max(order_by = Importance, n = 10, with_ties = FALSE) %>%
  mutate(
    Importance_scaled = Importance / max(Importance)
  ) %>%
  ungroup()

base_logit_features_plot <- base_logit_features_df %>%
  mutate(
    term_plot = tidytext::reorder_within(term_clean, Importance_scaled, Model_Label),
    Direction = factor(Direction, levels = c("Positive", "Negative"))
  )

ggplot(base_logit_features_plot, aes(x = term_plot, y = Importance_scaled, fill = Direction)) +
  geom_col(width = 0.7) +
  coord_flip() +
  tidytext::scale_x_reordered() +
  facet_wrap(~ Model_Label, ncol = 2, scales = "free_y") +
  theme_minimal(base_size = 12) +
  labs(
    title = "Top 10 Features for Base-Only Logistic-Family Models",
    x = NULL,
    y = "Relative Importance (scaled within model)",
    fill = "Effect"
  ) +
  theme(
    axis.text.y = element_text(size = 9),
    strip.text = element_text(size = 11, face = "bold"),
    legend.position = "bottom",
    panel.spacing = unit(1.2, "lines")
  )
```


# 4.2 Plot Top 10 Features for Base-Only Random Forest Model

```{r}
base_rf_features_df <- base_importance_all %>%
  filter(Model == "Random Forest") %>%
  group_by(Model_Label) %>%
  slice_max(order_by = Importance, n = 10, with_ties = FALSE) %>%
  mutate(
    Importance_scaled = Importance / max(Importance)
  ) %>%
  ungroup()

base_rf_features_plot <- base_rf_features_df %>%
  mutate(
    term_plot = tidytext::reorder_within(term_clean, Importance_scaled, Model_Label),
    Direction = factor(Direction, levels = c("RF Importance"))
  )

ggplot(base_rf_features_plot, aes(x = term_plot, y = Importance_scaled, fill = Direction)) +
  geom_col(width = 0.7) +
  coord_flip() +
  tidytext::scale_x_reordered() +
  facet_wrap(~ Model_Label, ncol = 1, scales = "free_y") +
  theme_minimal(base_size = 12) +
  labs(
    title = "Top 10 Features for Base-Only Random Forest",
    x = NULL,
    y = "Relative Importance (scaled within model)",
    fill = "Effect"
  ) +
  theme(
    axis.text.y = element_text(size = 9),
    strip.text = element_text(size = 11, face = "bold"),
    legend.position = "bottom",
    panel.spacing = unit(1.2, "lines")
  )
```


# 5. Fit Logistic Sentiment-Addons

## Why Add Sentiment Features?

While structured features (ratings, flags, word counts) capture explicit signals in employee reviews, they may miss the underlying tone and nuance in textual feedback. Sentiment analysis helps quantify this tone by transforming free-text fields into numerical features that reflect positivity, negativity, and variability across different parts of a review.

We incorporated three sentiment models:

- VADER: A rule-based sentiment analyzer designed for short, informal text. It produces a compound score capturing overall sentiment intensity.
- DistilBERT (distilbert-base-uncased-finetuned-sst-2-english): A lightweight transformer model fine-tuned for sentiment classification, providing strong contextual understanding with lower computational cost.
- RoBERTa (cardiffnlp/twitter-roberta-base-sentiment-latest): A more powerful transformer model trained on large-scale data, capable of capturing subtle linguistic nuances and contextual sentiment shifts.

By adding these features, we aim to test whether textual tone provides additional predictive power beyond structured variables.

```{r}
logit_results <- list()
logit_importance <- list()

for (name in names(datasets)) {
  df_current <- datasets[[name]]
  
  train_df <- df_current[train_index, ]
  test_df  <- df_current[-train_index, ]
  
  x_train <- model.matrix(flags_recommend_frend ~ ., data = train_df)[, -1]
  x_test  <- model.matrix(flags_recommend_frend ~ ., data = test_df)[, -1]
  y_train <- ifelse(train_df$flags_recommend_frend == "POSITIVE", 1, 0)
  
  model_logit <- glm(
    flags_recommend_frend ~ .,
    data = train_df,
    family = binomial
  )
  prob_logit <- predict(model_logit, newdata = test_df, type = "response")
  eval_logit <- evaluate_model(prob_logit, test_df$flags_recommend_frend)
  
  cv_lasso <- cv.glmnet(x_train, y_train, family = "binomial", alpha = 1)
  prob_lasso <- as.numeric(
    predict(cv_lasso, newx = x_test, s = "lambda.min", type = "response")
  )
  eval_lasso <- evaluate_model(prob_lasso, test_df$flags_recommend_frend)
  
  cv_ridge <- cv.glmnet(x_train, y_train, family = "binomial", alpha = 0)
  prob_ridge <- as.numeric(
    predict(cv_ridge, newx = x_test, s = "lambda.min", type = "response")
  )
  eval_ridge <- evaluate_model(prob_ridge, test_df$flags_recommend_frend)
  
  cv_enet <- cv.glmnet(x_train, y_train, family = "binomial", alpha = 0.5)
  prob_enet <- as.numeric(
    predict(cv_enet, newx = x_test, s = "lambda.min", type = "response")
  )
  eval_enet <- evaluate_model(prob_enet, test_df$flags_recommend_frend)
  
  logit_results[[name]] <- list(
    logit_model = model_logit,
    logit_eval = eval_logit,
    lasso_model = cv_lasso,
    lasso_eval = eval_lasso,
    ridge_model = cv_ridge,
    ridge_eval = eval_ridge,
    enet_model = cv_enet,
    enet_eval = eval_enet
  )
  
  logit_importance[[paste0(name, "_logit")]] <- extract_glm_importance(
    model_logit, name, "Logistic Regression"
  )
  logit_importance[[paste0(name, "_lasso")]] <- extract_glmnet_importance(
    cv_lasso, name, "LASSO LogReg"
  )
  logit_importance[[paste0(name, "_ridge")]] <- extract_glmnet_importance(
    cv_ridge, name, "Ridge LogReg"
  )
  logit_importance[[paste0(name, "_enet")]] <- extract_glmnet_importance(
    cv_enet, name, "Elastic Net LogReg"
  )
}
```

# 5.1 Summarize Logistic Models with Sentiment Add-Ons
```{r}
model_comparison <- do.call(rbind, lapply(names(logit_results), function(name) {
  res <- logit_results[[name]]
  
  data.frame(
    Features = name,
    Model = c("Logistic Regression", "LASSO LogReg", "Ridge LogReg", "Elastic Net LogReg"),
    AUC = c(
      res$logit_eval$auc,
      res$lasso_eval$auc,
      res$ridge_eval$auc,
      res$enet_eval$auc
    ),
    F1 = c(
      max(res$logit_eval$f1_results$F1),
      max(res$lasso_eval$f1_results$F1),
      max(res$ridge_eval$f1_results$F1),
      max(res$enet_eval$f1_results$F1)
    ),
    Best_Threshold = c(
      res$logit_eval$best_threshold,
      res$lasso_eval$best_threshold,
      res$ridge_eval$best_threshold,
      res$enet_eval$best_threshold
    ),
    Accuracy = c(
      as.numeric(res$logit_eval$confusion_matrix$overall["Accuracy"]),
      as.numeric(res$lasso_eval$confusion_matrix$overall["Accuracy"]),
      as.numeric(res$ridge_eval$confusion_matrix$overall["Accuracy"]),
      as.numeric(res$enet_eval$confusion_matrix$overall["Accuracy"])
    )
  )
}))

model_comparison %>% arrange(desc(AUC))
```

# 5.2 Plot Logistic Sentiment Add-On Models
```{r}
logistic_rank_plot <- model_comparison %>%
  mutate(
    Model_Label = paste(Features, Model, sep = " + "),
    AUC_Rank = dense_rank(desc(AUC)),
    F1_Rank = dense_rank(desc(F1)),
    Avg_Rank = (AUC_Rank + F1_Rank) / 2
  ) %>%
  arrange(Avg_Rank) %>%
  mutate(Model_Label = factor(Model_Label, levels = rev(Model_Label)))

ggplot(logistic_rank_plot, aes(y = Model_Label)) +
  geom_segment(
    aes(x = AUC_Rank, xend = F1_Rank, yend = Model_Label),
    color = "gray70",
    linewidth = 1
  ) +
  geom_point(aes(x = AUC_Rank, color = "AUC Rank"), size = 2.8) +
  geom_point(aes(x = F1_Rank, color = "F1 Rank"), size = 2.8) +
  geom_point(aes(x = Avg_Rank, color = "Average Rank"), size = 3.2, shape = 18) +
  scale_x_reverse() +
  theme_minimal() +
  labs(
    title = "Logistic Models with Sentiment Add-Ons: AUC Rank vs F1 Rank vs Average Rank",
    x = "Rank (1 = Best)",
    y = NULL,
    color = "Metric"
  )
```

# 6. Fit Random Forest Models with Sentiment Add-Ons
```{r}
rf_results <- list()
rf_importance <- list()

for (name in names(datasets)) {
  df_current <- datasets[[name]]
  
  train_df <- df_current[train_index, ]
  test_df  <- df_current[-train_index, ]
  
  p <- ncol(train_df) - 1
  
  rf_model <- train(
    flags_recommend_frend ~ .,
    data = train_df,
    method = "rf",
    trControl = trainControl(method = "cv", number = 10),
    tuneGrid = expand.grid(mtry = 1:p),
    ntree = 500,
    importance = TRUE
  )
  
  bestTune <- rf_model$bestTune
  
  rf_model_tuned <- train(
    flags_recommend_frend ~ .,
    data = train_df,
    method = "rf",
    trControl = trainControl(
      method = "cv",
      number = 10,
      classProbs = TRUE,
      summaryFunction = twoClassSummary
    ),
    tuneGrid = bestTune,
    ntree = 500,
    metric = "ROC",
    importance = TRUE
  )
  
  prob_rf <- predict(rf_model_tuned, newdata = test_df, type = "prob")[, "POSITIVE"]
  eval_rf <- evaluate_model(prob_rf, test_df$flags_recommend_frend)
  
  rf_results[[name]] <- list(
    best_mtry = bestTune$mtry,
    rf_model = rf_model,
    rf_model_tuned = rf_model_tuned,
    eval = eval_rf
  )
  
  rf_importance[[paste0(name, "_rf")]] <- extract_rf_importance(
    rf_model_tuned, name, "Random Forest"
  )
}
```

# 6.1 Summarize Random Forest Models with Sentiment Add-Ons
```{r}
rf_summary <- data.frame(
  Features = names(rf_results),
  Model = "Random Forest",
  Best_mtry = sapply(rf_results, function(x) x$best_mtry),
  AUC = sapply(rf_results, function(x) x$eval$auc),
  F1 = sapply(rf_results, function(x) max(x$eval$f1_results$F1)),
  Accuracy = sapply(rf_results, function(x) as.numeric(x$eval$confusion_matrix$overall["Accuracy"])),
  Best_Threshold = sapply(rf_results, function(x) x$eval$best_threshold),
  row.names = NULL
)

rf_summary
```

# 6.2 Plot Random Forest Sentiment Add-On Models
```{r}
rf_rank_plot <- rf_summary %>%
  mutate(
    Model_Label = paste(Features, Model, sep = " + "),
    AUC_Rank = dense_rank(desc(AUC)),
    F1_Rank = dense_rank(desc(F1)),
    Avg_Rank = (AUC_Rank + F1_Rank) / 2
  ) %>%
  arrange(Avg_Rank) %>%
  mutate(Model_Label = factor(Model_Label, levels = rev(Model_Label)))

ggplot(rf_rank_plot, aes(y = Model_Label)) +
  geom_segment(
    aes(x = AUC_Rank, xend = F1_Rank, yend = Model_Label),
    color = "gray70",
    linewidth = 1
  ) +
  geom_point(aes(x = AUC_Rank, color = "AUC Rank"), size = 3) +
  geom_point(aes(x = F1_Rank, color = "F1 Rank"), size = 3) +
  geom_point(aes(x = Avg_Rank, color = "Average Rank"), size = 3.5, shape = 18) +
  scale_x_reverse() +
  theme_minimal() +
  labs(
    title = "Random Forest Models with Sentiment Add-Ons: AUC Rank vs F1 Rank vs Average Rank",
    x = "Rank (1 = Best)",
    y = NULL,
    color = "Metric"
  )
```

# 7. Combine All Model Results
```{r}
logistic_df <- model_comparison %>%
  dplyr::select(Features, Model, AUC, F1, Accuracy)

rf_df <- rf_summary %>%
  dplyr::select(Features, Model, AUC, F1, Accuracy)

full_summary_df <- bind_rows(logistic_df, rf_df) %>%
  mutate(
    Model_Label = paste(Features, Model, sep = " + "),
    AUC_Rank = dense_rank(desc(AUC)),
    F1_Rank = dense_rank(desc(F1)),
    Avg_Rank = (AUC_Rank + F1_Rank) / 2
  ) %>%
  arrange(Avg_Rank)

full_summary_df
```



# 7.1 Plot Overall Model Rankings
```{r}
full_summary_plot <- full_summary_df %>%
  mutate(
    Avg_Rank = (AUC_Rank + F1_Rank) / 2
  ) %>%
  arrange(Avg_Rank) %>%
  mutate(
    Model_Label = factor(Model_Label, levels = rev(Model_Label))
  )

ggplot(full_summary_plot, aes(y = Model_Label)) +
  # Line between AUC and F1
  geom_segment(
    aes(x = AUC_Rank, xend = F1_Rank, yend = Model_Label),
    color = "gray70",
    linewidth = 1
  ) +
  
  # AUC point
  geom_point(aes(x = AUC_Rank, color = "AUC"), size = 3) +
  
  # F1 point
  geom_point(aes(x = F1_Rank, color = "F1"), size = 3) +
  
  # Avg rank point (NEW)
  geom_point(aes(x = Avg_Rank, color = "Average"), size = 3.5, shape = 18) +
  
  scale_x_reverse() +
  scale_color_manual(values = c(
    "AUC" = "#E57373",
    "F1" = "#4DB6AC",
    "Average" = "#5C6BC0"
  )) +
  
  theme_minimal(base_size = 12) +
  labs(
    title = "All Models: AUC vs F1 vs Average Rank",
    x = "Rank (1 = Best)",
    y = NULL,
    color = "Metric"
  ) +
  theme(
    legend.position = "bottom"
  )
```



# 8. View Top 10 Models by Average Rank
```{r}
top_models <- full_summary_df %>%
  slice_head(n = 10)

top_models %>%
  dplyr::select(Model_Label, AUC, F1, Accuracy, AUC_Rank, F1_Rank, Avg_Rank)
```

# 8.1 Plot Top 10 Models by Average Rank
```{r}
top_models_plot <- top_models %>%
  arrange(Avg_Rank) %>%
  mutate(Model_Label = factor(Model_Label, levels = rev(Model_Label)))

ggplot(top_models_plot, aes(y = Model_Label)) +
  geom_segment(
    aes(x = AUC_Rank, xend = F1_Rank, yend = Model_Label),
    color = "gray70",
    linewidth = 1
  ) +
  geom_point(aes(x = AUC_Rank, color = "AUC Rank"), size = 3) +
  geom_point(aes(x = F1_Rank, color = "F1 Rank"), size = 3) +
  geom_point(aes(x = Avg_Rank, color = "Average Rank"), size = 3.5, shape = 18) +
  scale_x_reverse() +
  theme_minimal() +
  labs(
    title = "Top 10 Models: AUC Rank vs F1 Rank vs Average Rank",
    x = "Rank (1 = Best)",
    y = NULL,
    color = "Metric"
  )
```

# 9. Combine All Feature Importance Results
```{r}
importance_all <- bind_rows(
  bind_rows(logit_importance),
  bind_rows(rf_importance)
) %>%
  mutate(
    Model_Label = paste(Features, Model, sep = " + "),
    term_clean = clean_feature_names(term)
  )

importance_all
```

# 9.1 View Top 10 Features for the Best Overall Models
```{r}
top_model_features_df <- importance_all %>%
  semi_join(top_models, by = "Model_Label") %>%
  group_by(Model_Label) %>%
  slice_max(order_by = Importance, n = 10, with_ties = FALSE) %>%
  mutate(Importance_scaled = Importance / max(Importance)) %>%
  ungroup() %>%
  arrange(Model_Label, desc(Importance_scaled))

top_model_features_df
```

# 9.2 Plot Top 10 Features for the Best Logistic-Family Models

```{r plot-top-10-features-best-logit-models, fig.width=10, fig.height=10}
top_logit_models <- top_models %>%
  filter(Model != "Random Forest")

top_logit_features_df <- importance_all %>%
  semi_join(top_logit_models, by = "Model_Label") %>%
  filter(Model != "Random Forest") %>%
  group_by(Model_Label) %>%
  slice_max(order_by = Importance, n = 10, with_ties = FALSE) %>%
  mutate(
    Importance_scaled = Importance / max(Importance)
  ) %>%
  ungroup()

top_logit_features_plot <- top_logit_features_df %>%
  mutate(
    term_plot = tidytext::reorder_within(term_clean, Importance_scaled, Model_Label),
    Direction = factor(Direction, levels = c("Positive", "Negative"))
  )

ggplot(top_logit_features_plot, aes(x = term_plot, y = Importance_scaled, fill = Direction)) +
  geom_col(width = 0.7) +
  coord_flip() +
  tidytext::scale_x_reordered() +
  facet_wrap(~ Model_Label, ncol = 2, scales = "free_y") +
  theme_minimal(base_size = 12) +
  labs(
    title = "Top 10 Features for the Best Logistic-Family Models",
    x = NULL,
    y = "Relative Importance (scaled within model)",
    fill = "Effect"
  ) +
  theme(
    axis.text.y = element_text(size = 9),
    strip.text = element_text(size = 11, face = "bold"),
    legend.position = "bottom",
    panel.spacing = unit(1.2, "lines")
  )
```


# 9.3 Top 10 Features for Best Random Forest Model
```{r}
top_rf_models <- top_models %>%
  filter(Model == "Random Forest")

if (nrow(top_rf_models) == 0) {
  cat("No Random Forest models appear in the top models table.\n")
} else {
  top_rf_features_df <- importance_all %>%
    semi_join(top_rf_models, by = "Model_Label") %>%
    filter(Model == "Random Forest") %>%
    group_by(Model_Label) %>%
    slice_max(order_by = Importance, n = 10, with_ties = FALSE) %>%
    mutate(
      Importance_scaled = Importance / max(Importance, na.rm = TRUE)
    ) %>%
    ungroup()

  if (nrow(top_rf_features_df) == 0) {
    cat("No Random Forest feature importance rows available for plotting.\n")
  } else {
    top_rf_features_plot <- top_rf_features_df %>%
      mutate(
        term_plot = tidytext::reorder_within(term_clean, Importance_scaled, Model_Label),
        Direction = factor(Direction, levels = c("RF Importance"))
      )

    ggplot(top_rf_features_plot, aes(x = term_plot, y = Importance_scaled, fill = Direction)) +
      geom_col(width = 0.7) +
      coord_flip() +
      tidytext::scale_x_reordered() +
      facet_wrap(~ Model_Label, ncol = 1, scales = "free_y") +
      theme_minimal(base_size = 12) +
      labs(
        title = "Top 10 Features for the Best Random Forest Models",
        x = NULL,
        y = "Relative Importance (scaled within model)",
        fill = "Effect"
      ) +
      theme(
        axis.text.y = element_text(size = 9),
        strip.text = element_text(size = 11, face = "bold"),
        legend.position = "bottom",
        panel.spacing = unit(1.2, "lines")
      )
  }
}
```


