---
title: "modeling-core-mahin"
author: "Mahin Roddur"
date: "2026-04-08"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(caret)
library(pROC)
library(MASS)       # polr
library(broom)
library(cluster)    # silhouette
library(factoextra) # elbow / clustering plots
library(glmnet)     # optional regularization
```

```{r}
library(here)
data <- read.csv(here("data", "processed", "reviews_with_sent_scores.csv"))
```

```{r}
data <- data %>%
  mutate(
    employee_length_bin = case_when(
      employee_length <= 1 ~ "short",
      employee_length < 4 ~ "medium",
      TRUE ~ "long"
    ),
    employee_length_bin = factor(
      employee_length_bin,
      levels = c("short", "medium", "long"),
      ordered = TRUE
    ))
```

```{r}
ggplot(data, aes(x = employee_length_bin)) +
  geom_bar() +
  theme_minimal() +
  labs(title = "Distribution of Employee Tenure Categories")
```
The bins are not heavily imbalanced. Yet, it was hard to predict, as we will see.

```{r}

features_base <- c(
  # Ratings
  "rating_culture_values",
  "rating_work_life",
  "rating_career_opportunities",
  "rating_compensation_benefits",
  "rating_senior_leadership",
  "rating_diversity_inclusion",
  
  # Context
  "employee_job_quit",
  "covid_related_review",
  "rating_year",
  
  # Missingness
  "career_missing",
  
  # Engagement
  "helpfulness_score",
  
  # Text structure
  "summary_wc",
  "pros_wc",
  "cons_wc",
  "advice_wc"
)
```

Scaling helper
```{r}
scale_train_test <- function(train_df, test_df, feature_cols) {
  train_scaled <- train_df
  test_scaled  <- test_df
  
  train_means <- sapply(train_df[feature_cols], mean, na.rm = TRUE)
  train_sds   <- sapply(train_df[feature_cols], sd, na.rm = TRUE)
  
  train_sds[train_sds == 0 | is.na(train_sds)] <- 1
  
  train_scaled[feature_cols] <- scale(train_df[feature_cols],
                                      center = train_means,
                                      scale = train_sds)
  
  test_scaled[feature_cols] <- scale(test_df[feature_cols],
                                     center = train_means,
                                     scale = train_sds)
  
  list(
    train = train_scaled,
    test = test_scaled,
    center = train_means,
    scale = train_sds
  )
}
```

# Train
```{r}
split_data <- function(df, target, p = 0.8, seed = 123) {
  set.seed(seed)
  idx <- createDataPartition(df[[target]], p = p, list = FALSE)
  list(
    train = df[idx, ],
    test  = df[-idx, ]
  )
}

rmse_vec <- function(actual, predicted) {
  sqrt(mean((actual - predicted)^2, na.rm = TRUE))
}

```

```{r}
run_tenure_lm <- function(df, feature_cols, target = "employee_length", seed = 123) {
  df_model <- data %>%
  dplyr::select(employee_length, all_of(features_base))
  
  split_obj <- split_data(df_model, target, seed = seed)
  train_df <- split_obj$train
  test_df  <- split_obj$test
  
  scaled <- scale_train_test(train_df, test_df, feature_cols)
  train_scaled <- scaled$train
  test_scaled  <- scaled$test
  
  formula_lm <- as.formula(
    paste(target, "~", paste(feature_cols, collapse = " + "))
  )
  
  model <- lm(formula_lm, data = train_scaled)
  
  preds <- predict(model, newdata = test_scaled)
  rmse  <- rmse_vec(test_scaled[[target]], preds)
  
  coef_tbl <- broom::tidy(model) %>%
    filter(term != "(Intercept)") %>%
    arrange(desc(abs(estimate)))
  
  list(
    model = model,
    rmse = rmse,
    predictions = preds,
    actual = test_scaled[[target]],
    coefficients = coef_tbl,
    train = train_scaled,
    test = test_scaled
  )
}

tenure_lm_results <- run_tenure_lm(
  df = data,
  feature_cols = features_base,
  target = "employee_length"
)

tenure_lm_results$rmse
tenure_lm_results$coefficients
summary(tenure_lm_results$model)
```
Abysmal R-squared 0.03894. RMSE is low but the model is barely better than randomly guessing.

```{r}
df_plot <- data.frame(
  actual = tenure_lm_results$actual,
  predicted = tenure_lm_results$predictions
)

df_plot$error <- df_plot$predicted - df_plot$actual

ggplot(df_plot, aes(x = actual, y = predicted, color = error)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  scale_color_gradient2(low = "blue", mid = "white", high = "red") +
  theme_minimal()
```
Looks goofy, making employee_length into ordinal bins might help. The model systematically underestimates long-tenure employees and overestimates short-tenure employees.

Really low adjusted R-squared: 0.03894. Attempting LASSO but doubt it will help.

```{r}
library(glmnet)

run_lasso_tenure <- function(df, feature_cols, target = "employee_length", seed = 123) {
  df_model <- data %>%
  dplyr::select(employee_length, all_of(features_base))
  
  split_obj <- split_data(df_model, target, seed = seed)
  train_df <- split_obj$train
  test_df  <- split_obj$test
  
  scaled <- scale_train_test(train_df, test_df, feature_cols)
  train_scaled <- scaled$train
  test_scaled  <- scaled$test
  
  x_train <- model.matrix(
    as.formula(paste(target, "~", paste(feature_cols, collapse = " + "))),
    data = train_scaled
  )[ , -1, drop = FALSE]
  
  y_train <- train_scaled[[target]]
  
  x_test <- model.matrix(
    as.formula(paste(target, "~", paste(feature_cols, collapse = " + "))),
    data = test_scaled
  )[ , -1, drop = FALSE]
  
  y_test <- test_scaled[[target]]
  
  cv_model <- cv.glmnet(
    x = x_train,
    y = y_train,
    alpha = 1,
    family = "gaussian"
  )
  
  best_lambda <- cv_model$lambda.min
  
  preds <- as.numeric(predict(cv_model, s = best_lambda, newx = x_test))
  rmse  <- rmse_vec(y_test, preds)
  
  coef_tbl <- as.matrix(coef(cv_model, s = "lambda.min")) %>%
    as.data.frame()
  
  colnames(coef_tbl) <- "estimate"
  coef_tbl$term <- rownames(coef_tbl)
  
  coef_tbl <- coef_tbl %>%
    dplyr::filter(term != "(Intercept)", estimate != 0) %>%
    dplyr::arrange(desc(abs(estimate)))
  
  list(
    model = cv_model,
    lambda = best_lambda,
    rmse = rmse,
    predictions = preds,
    actual = y_test,
    coefficients = coef_tbl,
    train = train_scaled,
    test = test_scaled
  )
}

tenure_lasso_results <- run_lasso_tenure(
  df = data,
  feature_cols = features_base,
  target = "employee_length"
)

tenure_lasso_results$rmse
tenure_lasso_results$coefficients
plot(tenure_lasso_results$model)
```
Regularization is cooking without the stove on. MSE barely changes across lambda. Give it up.

# Ordinal Regression (polr) for employee_length
```{r}
run_tenure_ordinal <- function(df, feature_cols, target = "length_bin", seed = 123) {
  df_model <- data %>%
  dplyr::select(employee_length_bin, all_of(features_base))
  
  split_obj <- split_data(df_model, target, seed = seed)
  train_df <- split_obj$train
  test_df  <- split_obj$test
  
  scaled <- scale_train_test(train_df, test_df, feature_cols)
  train_scaled <- scaled$train
  test_scaled  <- scaled$test
  
  formula_ord <- as.formula(
    paste(target, "~", paste(feature_cols, collapse = " + "))
  )
  
  model <- polr(formula_ord, data = train_scaled, Hess = TRUE)
  
  prob_preds <- predict(model, newdata = test_scaled, type = "probs")
  class_preds <- predict(model, newdata = test_scaled, type = "class")
  
  acc <- mean(as.character(class_preds) == as.character(test_scaled[[target]]), na.rm = TRUE)
  
  coef_tbl <- broom::tidy(model) %>%
    filter(!grepl("\\|", term)) %>%
    arrange(desc(abs(estimate)))
  
  list(
    model = model,
    accuracy = acc,
    class_predictions = class_preds,
    actual = test_scaled[[target]],
    coefficients = coef_tbl,
    train = train_scaled,
    test = test_scaled
  )
}

tenure_ord_results <- run_tenure_ordinal(
  df = data,
  feature_cols = features_base,
  target = "employee_length_bin"
)

tenure_ord_results$accuracy
tenure_ord_results$coefficients
summary(tenure_ord_results$model)
```
40% accuracy still pretty bad, better than random (0.33). Huge variance in the bins (1 vs 20  years). Employee tenure might be just hard to predict with the variables available in the dataset. 

Does NLP on the text fields help?

# Logistic Regression on flag_recommend_frend

```{r}
run_recommend_logit <- function(df, feature_cols, target = "flags_recommend_frend", seed = 123) {
  df_model <- data %>%
  dplyr::select(flags_recommend_frend, all_of(features_base))

  df_model[[target]] <- factor(df_model[[target]], levels = c("NEGATIVE", "POSITIVE"))

  split_obj <- split_data(df_model, target, seed = seed)
  train_df <- split_obj$train
  test_df  <- split_obj$test

  scaled <- scale_train_test(train_df, test_df, feature_cols)
  train_scaled <- scaled$train
  test_scaled  <- scaled$test

  train_scaled[[target]] <- factor(train_scaled[[target]], levels = c("NEGATIVE", "POSITIVE"))
  test_scaled[[target]]  <- factor(test_scaled[[target]], levels = c("NEGATIVE", "POSITIVE"))

  formula_logit <- as.formula(
    paste(target, "~", paste(feature_cols, collapse = " + "))
  )

  model <- glm(formula_logit, data = train_scaled, family = binomial)

  prob_preds <- predict(model, newdata = test_scaled, type = "response")

  class_preds <- ifelse(prob_preds >= 0.5, "POSITIVE", "NEGATIVE")
  class_preds <- factor(class_preds, levels = levels(test_scaled[[target]]))

  acc <- mean(class_preds == test_scaled[[target]], na.rm = TRUE)

  roc_obj <- pROC::roc(
    response = test_scaled[[target]],
    predictor = prob_preds,
    levels = c("NEGATIVE", "POSITIVE"),
    direction = "<"
  )
  auc_val <- as.numeric(pROC::auc(roc_obj))

  coef_tbl <- broom::tidy(model) %>%
    dplyr::filter(term != "(Intercept)") %>%
    dplyr::mutate(odds_ratio = exp(estimate)) %>%
    dplyr::arrange(desc(abs(estimate)))

  list(
    model = model,
    auc = auc_val,
    accuracy = acc,
    roc = roc_obj,
    probabilities = prob_preds,
    predicted_class = class_preds,
    actual = test_scaled[[target]],
    coefficients = coef_tbl,
    train = train_scaled,
    test = test_scaled
  )
}

recommend_results <- run_recommend_logit(
  df = data,
  feature_cols = features_base,
  target = "flags_recommend_frend"
)

recommend_results$auc
recommend_results$accuracy
recommend_results$coefficients
summary(recommend_results$model)
```
Best result yet. The model achieved an accuracy of approximately 0.864 and an AUC of 0.911, indicating strong predictive performance.

In contrast to the tenure models, recommendation behavior is well explained by workplace ratings, particularly culture, leadership, and work-life balance.

Reason for including AUC: AUC is particularly useful because it incorporates the full range of predicted probabilities and is less sensitive to threshold choice, making it a more robust measure of overall model quality.

## Ranking Features from the recommend logit model

```{r}
get_ranked_coefs <- function(model) {
  
  coefs <- coef(model)
  
  df <- data.frame(
    feature = names(coefs),
    value = as.numeric(coefs)
  )
  
  df <- df[df$feature != "(Intercept)", ]
  
  df <- df %>%
    mutate(abs_value = abs(value)) %>%
    arrange(desc(abs_value))
  
  df
}

coef_df <- get_ranked_coefs(recommend_results$model)
head(coef_df, 10)

```

```{r}
ggplot(coef_df[1:10, ], aes(x = reorder(feature, abs_value), y = value)) +
  geom_col() +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Top Features Driving Recommendation",
    x = "Feature",
    y = "Coefficient"
  )
```
Rating variables rank the highest. Interestingly, helpfulness score is strongly negatively correlated with recommendation - more helpful reviews tend to be more critical (not likely to be recommended).

## Sidebar: Rating correlation matrix
```{r}
rating_features <- c(
  "rating_culture_values",
  "rating_work_life",
  "rating_career_opportunities",
  "rating_compensation_benefits",
  "rating_senior_leadership",
  "rating_diversity_inclusion"
)

cor_mat <- cor(data[rating_features], use = "pairwise.complete.obs")
library(ggcorrplot)

cor_mat <- cor(data[rating_features], use = "pairwise.complete.obs")

ggcorrplot(
  cor_mat,
  type = "lower",
  lab = TRUE,
  lab_size = 3,
  colors = c("blue", "white", "red"),
  title = "Correlation Between Rating Dimensions"
) +
  theme_minimal() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```
Correlation analysis shows that workplace rating variables are moderately correlated, indicating the presence of multicollinearity. As a result, coefficient estimates should be interpreted as partial effects rather than independent contributions of each feature.



```{r}
plot_vif <- function(model, title = "VIF Plot") {
  library(car)
  library(dplyr)
  library(ggplot2)
  
  vif_vals <- vif(model)
  
  vif_df <- data.frame(
    feature = names(vif_vals),
    vif = as.numeric(vif_vals)
  ) %>%
    arrange(desc(vif))
  
  ggplot(vif_df, aes(x = reorder(feature, vif), y = vif)) +
    geom_col(fill = "steelblue") +
    coord_flip() + 
    geom_hline(yintercept = 5, linetype = "dashed", color = "red") +
    geom_hline(yintercept = 2, linetype = "dotted", color = "orange") +
    theme_minimal() +
    labs(
      title = title,
      x = "Feature",
      y = "VIF"
    ) 
}

plot_vif(tenure_lm_results$model, "VIF - Employee Length Model")
plot_vif(recommend_results$model, "VIF - Recommendation Model")

```
The employee_length model shows moderate multicollinearity among workplace rating variables, with VIF values around 3–3.5 for culture and senior leadership. This reflects the inherent correlation between different aspects of workplace evaluation. However, all VIF values remain below common thresholds of concern (VIF > 5), indicating that multicollinearity is present but not severe.

The recommendation model exhibits low VIF values (all below 2.5), indicating minimal multicollinearity. This suggests that the estimated effects of workplace features are stable and can be interpreted with confidence.

# Regularization attempts with LASSO, Ridge and Elastic Net on recommendation model (even though multicollinearity is not that severe)
```{r}
run_glmnet_logit <- function(df, feature_cols, target = "flags_recommend_frend", alpha = 1, seed = 123) {
  df_model <- df %>%
    dplyr::select(dplyr::all_of(c(target, feature_cols))) %>%
    dplyr::filter(!is.na(.data[[target]]))
  
  df_model[[target]] <- factor(df_model[[target]], levels = c("NEGATIVE", "POSITIVE"))
  
  split_obj <- split_data(df_model, target, seed = seed)
  train_df <- split_obj$train
  test_df  <- split_obj$test
  
  scaled <- scale_train_test(train_df, test_df, feature_cols)
  train_scaled <- scaled$train
  test_scaled  <- scaled$test
  
  train_scaled[[target]] <- factor(train_scaled[[target]], levels = c("NEGATIVE", "POSITIVE"))
  test_scaled[[target]]  <- factor(test_scaled[[target]], levels = c("NEGATIVE", "POSITIVE"))
  
  x_train <- model.matrix(
    as.formula(paste(target, "~", paste(feature_cols, collapse = " + "))),
    data = train_scaled
  )[ , -1, drop = FALSE]
  
  x_test <- model.matrix(
    as.formula(paste(target, "~", paste(feature_cols, collapse = " + "))),
    data = test_scaled
  )[ , -1, drop = FALSE]
  
  y_train <- ifelse(train_scaled[[target]] == "POSITIVE", 1, 0)
  y_test  <- ifelse(test_scaled[[target]] == "POSITIVE", 1, 0)
  
  set.seed(seed)
  cv_fit <- glmnet::cv.glmnet(
    x_train, y_train,
    family = "binomial",
    alpha = alpha
  )
  
  prob_preds <- as.numeric(
    predict(cv_fit, s = "lambda.min", newx = x_test, type = "response")
  )
  
  class_preds <- ifelse(prob_preds >= 0.5, 1, 0)
  acc <- mean(class_preds == y_test, na.rm = TRUE)
  
  roc_obj <- pROC::roc(response = y_test, predictor = prob_preds)
  auc_val <- as.numeric(pROC::auc(roc_obj))
  
  coef_tbl <- as.matrix(coef(cv_fit, s = "lambda.min")) %>%
    as.data.frame()
  colnames(coef_tbl) <- "estimate"
  coef_tbl$term <- rownames(coef_tbl)
  
  coef_tbl <- coef_tbl %>%
    dplyr::filter(term != "(Intercept)", estimate != 0) %>%
    dplyr::arrange(desc(abs(estimate)))
  
  list(
    cv_model = cv_fit,
    lambda_min = cv_fit$lambda.min,
    lambda_1se = cv_fit$lambda.1se,
    auc = auc_val,
    accuracy = acc,
    coefficients = coef_tbl,
    probabilities = prob_preds,
    actual = y_test
  )
}

lasso_rec <- run_glmnet_logit(data, features_base, alpha = 1)
ridge_rec <- run_glmnet_logit(data, features_base, alpha = 0)
enet_rec  <- run_glmnet_logit(data, features_base, alpha = 0.5)

reg_compare <- data.frame(
  model = c("Logistic", "LASSO", "Ridge", "Elastic Net"),
  auc = c(
    recommend_results$auc,
    lasso_rec$auc,
    ridge_rec$auc,
    enet_rec$auc
  ),
  accuracy = c(
    recommend_results$accuracy,
    lasso_rec$accuracy,
    ridge_rec$accuracy,
    enet_rec$accuracy
  )
)

reg_compare
```
Regularization barely changed performance. Saving LASSO for sentiment add-on later on.

```{r}
#saveRDS(lasso_rec, "models/best_recommend_base_model.rds")
#saveRDS(tenure_ord_results, "models/employee_length_bin_model.rds")
```


# Cluster Chaos

Clustering Prep for review-level clustering, company-level would be (was) too weak

