---
title: "modeling-mahin"
output: html_document
---

```{r setup, include=FALSE}
```

```{r}
library(here)
data <- read.csv(here("data", "processed", "reviews_with_sent_scores.csv"))
```


# Checking imbalance in response variables

```{r}
table(data$flags_recommend_frend)
prop.table(table(data$flags_recommend_frend))
```
```{r}
table(data$rating_overall)
prop.table(table(data$rating_overall))
```
```{r}
table(data$employee_length)
summary(data$employee_length)
```
Employee length seems to be a good candidate to bin into ordinal categories, ie. short, medium, long.

```{r}

data <- data %>%
  mutate(
    employee_length_bin = case_when(
      employee_length <= 1 ~ "short",
      employee_length < 4 ~ "medium",
      TRUE ~ "long"
    ),
    employee_length_bin = factor(
      employee_length_bin,
      levels = c("short", "medium", "long"),
      ordered = TRUE
    ))
table(data$employee_length_bin)
```


```{r}
data$sent_vader_avg <- rowMeans(
  data[, c(
    "vader_summary_compound",
    "vader_pros_compound",
    "vader_cons_compound",
    "vader_advice_compound"
  )],
  na.rm = TRUE
)

data$sent_vader_sd <- apply(
  data[, c("vader_summary_compound",
    "vader_pros_compound",
    "vader_cons_compound",
    "vader_advice_compound")],
  1,
  sd,
  na.rm = TRUE
)

data$sent_distilbert_avg <- rowMeans(
  data[, c(
    "distilbert_summary_signed",
    "distilbert_pros_signed",
    "distilbert_cons_signed",
    "distilbert_advice_signed"
  )],
  na.rm = TRUE
)

data$sent_distilbert_sd <- apply(
  data[, c("distilbert_summary_signed",
         "distilbert_pros_signed",
         "distilbert_cons_signed",
         "distilbert_advice_signed")],
  1,
  sd,
  na.rm = TRUE
)

data$sent_roberta_avg <- rowMeans(
  data[, c(
    "roberta_summary_signed",
    "roberta_pros_signed",
    "roberta_cons_signed",
    "roberta_advice_signed"
  )],
  na.rm = TRUE
)

data$sent_roberta_sd <- apply(
  data[, c(    "roberta_summary_signed",
    "roberta_pros_signed",
    "roberta_cons_signed",
    "roberta_advice_signed")],
  1,
  sd,
  na.rm = TRUE
)
```


```{r}
# =========================
# 1) Core structured features
# =========================

features_base <- c(
  # Ratings
  "rating_culture_values",
  "rating_work_life",
  "rating_career_opportunities",
  "rating_compensation_benefits",
  "rating_senior_leadership",
  "rating_diversity_inclusion",
  
  # Context
  "employee_job_quit",
  "covid_related_review",
  "rating_year",
  
  # Missingness
  "career_missing",
  
  # Engagement
  "helpfulness_score",
  
  # Text structure
  "summary_wc",
  "pros_wc",
  "cons_wc",
  "advice_wc"
)

# =========================
# 2) Sentiment feature sets by model
# =========================
# avg = overall tone across summary/pros/cons/advice
# sd  = disagreement / variation across review sections

features_vader <- c(
  features_base,
  "sent_vader_avg",
  "sent_vader_sd"
)

features_distilbert <- c(
  features_base,
  "sent_distilbert_avg",
  "sent_distilbert_sd"
)

features_roberta <- c(
  features_base,
  "sent_roberta_avg",
  "sent_roberta_sd"
)

feature_sets_avg_only <- list(
  structured_only = features_base,
  vader = c(features_base, "sent_vader_avg"),
  distilbert = c(features_base, "sent_distilbert_avg"),
  roberta = c(features_base, "sent_roberta_avg")
)

# =========================
# 3) Optional baseline without sentiment
# =========================
# Use this to test whether sentiment adds predictive value at all

features_structured_only <- features_base

# =========================
# 4) Bundle all feature sets for looping
# =========================

feature_sets <- list(
  structured_only = features_structured_only,
  vader = features_vader,
  distilbert = features_distilbert,
  roberta = features_roberta
)


```

# Baseline  vs Sentiment modeling

```{r}
library(dplyr)

df_model <- data %>%
  rename(flags_recommend_friend = flags_recommend_frend) %>%   # fix typo in original column name
  mutate(
    # binary classification target
    flags_recommend_friend = factor(
      flags_recommend_friend,
      levels = c("NEGATIVE", "POSITIVE")
    ),
    
    # ordinal target
    length_bin = ordered(
      length_bin,
      levels = c("short", "medium", "long")
    ),
    
    # convert logical predictors to numeric 0/1 for modeling
    employee_job_quit = as.integer(employee_job_quit),
    career_missing = as.integer(career_missing),
    covid_related_review = as.integer(covid_related_review)
  )
```

```{r}
df_reg <- df_model %>%
  filter(!is.na(rating_overall))

df_bin <- df_model %>%
  filter(!is.na(flags_recommend_friend))

df_ord <- df_model %>%
  filter(!is.na(length_bin))
```


# Training each model

```{r}
library(caret)
set.seed(123)

# Regression
idx_reg <- createDataPartition(df_reg$rating_overall, p = 0.8, list = FALSE)
train_reg <- df_reg[idx_reg, ]
test_reg  <- df_reg[-idx_reg, ]

# Binary classification
idx_bin <- createDataPartition(df_bin$flags_recommend_friend, p = 0.8, list = FALSE)
train_bin <- df_bin[idx_bin, ]
test_bin  <- df_bin[-idx_bin, ]

# Ordinal
idx_ord <- createDataPartition(df_ord$length_bin, p = 0.8, list = FALSE)
train_ord <- df_ord[idx_ord, ]
test_ord  <- df_ord[-idx_ord, ]
```

# Basic Linear Model Fitting on rating_overall
```{r}
results_list <- list()

for (name in names(feature_sets)) {
  
  features <- feature_sets[[name]]
  
  # Fit model
  model <- lm(
    rating_overall ~ .,
    data = train_reg[, c("rating_overall", features)]
  )
  
  # Predictions
  preds <- predict(model, newdata = test_reg)
  
  # RMSE
  rmse <- sqrt(mean((test_reg$rating_overall - preds)^2))
  
  # Extract coefficients
  coef_df <- as.data.frame(summary(model)$coefficients)
  coef_df$feature <- rownames(coef_df)
  coef_df <- coef_df[coef_df$feature != "(Intercept)", ]
  
  # sort by absolute effect
  coef_df <- coef_df[order(-abs(coef_df$Estimate)), ]
  
  # take top 5 features
  top_features <- head(coef_df$feature, 5)
  
  # collapse into a string
  top_features_str <- paste(top_features, collapse = ", ")
  
  # store results
  results_list[[name]] <- data.frame(
    model = name,
    RMSE = rmse,
    top_features = top_features_str
  )
}

results_reg <- do.call(rbind, results_list)
results_reg <- results_reg[order(results_reg$RMSE), ]
results_reg <- data.frame(results_reg, row.names = NULL)

results_reg
```
Across all models, senior leadership and company culture consistently emerged as the strongest predictors of overall ratings. 

The inclusion of sentiment features improved predictive performance modestly, with VADER achieving the lowest RMSE. This is surprising because transformers are generally better at sentiment inference than the simpler lexicon-based method.  Maybe this can be explained because of the generally short average word counts across all text-fields.

The same structured variables remained dominant even when sentiment features were added, indicating that sentiment provides complementary rather than substitutive information.


# Recommend Friend Baseline Classification

```{r}
library(pROC)
library(caret)

results_bin_list <- list()

for (name in names(feature_sets)) {
  
  features <- feature_sets[[name]]
  
  # Fit logistic model
  model <- glm(
    flags_recommend_friend ~ .,
    data = train_bin[, c("flags_recommend_friend", features)],
    family = "binomial"
  )
  
  # Predict probabilities
  probs <- predict(model, newdata = test_bin, type = "response")
  
  # Convert to class (threshold = 0.5)
  preds <- ifelse(probs > 0.5, "POSITIVE", "NEGATIVE")
  preds <- factor(preds, levels = c("NEGATIVE", "POSITIVE"))
  
  # Accuracy
  acc <- mean(preds == test_bin$flags_recommend_friend)
  
  # AUC
  auc <- roc(
    response = test_bin$flags_recommend_friend,
    predictor = probs,
    levels = c("NEGATIVE", "POSITIVE")
  )$auc
  
  # Store
  results_bin_list[[name]] <- data.frame(
    model = name,
    AUC = as.numeric(auc),
    Accuracy = acc
  )
  
  # Extract coefficients
  coef_df <- as.data.frame(summary(model)$coefficients)
  coef_df$feature <- rownames(coef_df)
  coef_df <- coef_df[coef_df$feature != "(Intercept)", ]
  
  # sort
  coef_df <- coef_df[order(-abs(coef_df$Estimate)), ]
  
  # top 5 with sign
  top_features <- head(
    paste0(
      coef_df$feature,
      " (", round(coef_df$Estimate, 2), ")"
    ),
    5
  )
  
  top_features_str <- paste(top_features, collapse = ", ")
}



results_bin <- do.call(rbind, results_bin_list)
rownames(results_bin) <- NULL

results_bin[order(-results_bin$AUC), ]
```
Structured workplace ratings alone are sufficient to predict recommendation behavior with very high accuracy. 
The addition of sentiment features provides only marginal improvements in AUC and does not improve classification accuracy.

We prioritize AUC over accuracy as the primary evaluation metric because AUC evaluates the model’s ability to correctly rank positive and negative instances across all classification thresholds, whereas accuracy depends on a fixed threshold (0.5) and can be misleading, especially in imbalanced datasets.


# Employee Length
```{r}
library(MASS)

results_ord_list <- list()

for (name in names(feature_sets)) {
  
  features <- feature_sets[[name]]
  
  model <- polr(
    length_bin ~ .,
    data = train_ord[, c("length_bin", features)],
    Hess = TRUE
  )
  
  # Predictions
  preds <- predict(model, newdata = test_ord)
  
  # Accuracy
  acc <- mean(as.character(preds) == as.character(test_ord$length_bin))
  
  results_ord_list[[name]] <- data.frame(
    model = name,
    Accuracy = acc
  )
}

results_ord <- do.call(rbind, results_ord_list)
rownames(results_ord) <- NULL

results_ord[order(-results_ord$Accuracy), ]
```
Prediction of employee length proved more challenging than other tasks, with model accuracy ranging from 0.43 to 0.46. However, all models outperformed the random baseline of approximately 0.33. 
Sentiment features provided modest improvements over structured features alone, with DistilBERT achieving the best performance. This suggests that more nuanced language representations may better capture signals related to employee tenure.


# Overall Picture

Sentiment plays a more meaningful role in predicting continuous satisfaction and tenure-related outcomes than in predicting recommendation behavior, where structured ratings already capture most of the signal.
