---
title: "Job Quit Factor Analysis"
author: "Daniel Cheung"
date: "2026-04-12"
output:
  pdf_document: default
  html_document: default
---

Main conclusion: Covid is the most significant protector from quitting the job; missing of career opportunity is the most significant factor leading to job quitting; leadership could negatively affect the willing to stay, but after working 3 years at a company, staffs are more rational, and less sensitive to ratings.

# Job Quit Factor Analysis


```{r, message = F, warning = F}
if (!requireNamespace("pacman", quietly = TRUE)) {
  install.packages("pacman")
}
library(pacman)

p_load(here)
p_load(tidyverse)
p_load(caret)

p_load(car) # VIF


p_load(survival) # Cox
p_load(survminer) # Plot

p_load(randomForestSRC)

data <- read.csv(here("data", "processed", "reviews_with_sent_scores.csv"))

```



```{r}
features_base <- c(
  # Ratings
  # "rating_culture_values",
  "rating_work_life",
  "rating_career_opportunities",
  "rating_compensation_benefits",
  "rating_senior_leadership",
  "rating_diversity_inclusion",
  
  # Context
  "covid_related_review",
  
  # Missingness
  "career_missing"

)


```


## Cox Model

```{r}
surv_obj <- Surv(time = data$employee_length, event = data$employee_job_quit)
plot(survfit(surv_obj ~ 1), main = "Job Retention Ratio by Employee Length", ylim = c(0,1))

```

This plot illustrates how likely employees are to stay in their position. For this dataset, more than half of the employees are likely to quit their job within 3 years, and the porportion of long-term staff is less than 10%.

### Model 1: whole dataset
```{r}
coxmodel1 <- coxph(surv_obj ~ ., data[,features_base])
summary(coxmodel1)

ggforest(coxmodel1, data = data, main = "Hazard Ratio")

```

The Cox Model is built to find factors employees quit the job. The Concordance index is 0.592, thus the model is not bad.

The career_missing are the most significant factor shown in the model, indicating when employee doesn't rate their career opportunity are more likely to quit the job.

Note: ```rating_culture_values``` is dropped since high correlation with ```rating_senior_leadership```, which made VIF > 3. After dropping ```rating_culture_values```, the C-Index wasn't decreased too much. 

Covid is a protector of employee leaving, during the covid period people are less likely to quit their job.


<!-- Senior leadership and the that people may have higher expection of their boss and  -->
### Model 2: subset with year < 3

```{r}
data2 <- data[data$employee_length<3,]
surv_obj2 <- Surv(time = data2$employee_length, event = data2$employee_job_quit)

coxmodel2 <- coxph(surv_obj2 ~ ., data2[, features_base])
summary(coxmodel2)

ggforest(coxmodel2, data = data2, main = "Hazard Ratio for Employee Length < 3yr")

```

### Model 3: subset with year 3~5

```{r}
data3 <- data[(data$employee_length<5) & (data$employee_length>=3),]
surv_obj3 <- Surv(time = data3$employee_length, event = data3$employee_job_quit)

coxmodel3 <- coxph(surv_obj3 ~ ., data3[, features_base])
summary(coxmodel3)

ggforest(coxmodel3, data = data3, main = "Hazard Ratio for Employee Length 3 ~ 5yr")

```


The initial observation that higher leadership ratings act as a risk factor is a paradox that dissolves when examining mid-tenure staff. During the 3–5 year window, the influence of leadership quality completely neutralizes (HR = 1.01, p=0.81). This suggests a clear maturation effect among employees. By their third year, staff have matured past the stage where subjective leadership styles dictate their retention.

Instead, the model is entirely dominated by the absence of career opportunities (Career Missing, HR=1.35). This demonstrates that for experienced professionals, attrition is no longer about dissatisfaction with management, but rather a pragmatic response to a stagnant career trajectory. If the company fails to provide clear promotion pathways within this 3–5 year window, even the most highly-rated leaders cannot prevent turnover.


