---
title: "RQ2_Brian"
author: "Brian Park"
date: "2026-04-09"
output:
  pdf_document: default
  html_document: default
---


# Libraries
```{r}
pkg_needed <- c("tidyverse", "cluster", "factoextra", "pheatmap", "ggrepel")
for (p in pkg_needed) {
  if (!requireNamespace(p, quietly = TRUE)) {
    install.packages(p)
  }
}

library(pheatmap)
library(tidyverse)
library(cluster)
library(factoextra)
library(here)
library(cluster)
library(ggrepel)
```

# For Cleaner Plots
```{r}
theme_set(
  theme_minimal(base_size = 13) +
    theme(
      plot.title = element_text(face = "bold", size = 15),
      plot.subtitle = element_text(size = 11),
      axis.title = element_text(face = "bold"),
      panel.grid.minor = element_blank(),
      legend.position = "bottom"
    )
)
```

# Read from current data version
```{r}
df <- read.csv(here("data", "processed", "glassdoor_reviews_cleaned_v2.csv"))

dim(df)
head(df)
```

# 1. Feature Engineering

### Variables that matter for culture similarity
```{r}
# Select variables related to company culture / employee experience
glassdoor_model <- df %>%
  select(
    company_name,
    rating_work_life,
    rating_culture_values,
    rating_senior_leadership,
    rating_career_opportunities,
    rating_compensation_benefits,
    rating_overall,
    rating_diversity_inclusion,
    flags_recommend_frend,
    flags_business_outlook,
    flags_ceo_approval,
    employee_status
  )

# Check missing values
colSums(is.na(glassdoor_model))
```

### Convert categorical sentiment variables into numeric form
```{r}
# Recode recommendation / outlook / CEO approval into numeric variables
glassdoor_model <- glassdoor_model %>%
  mutate(
    recommend_num = ifelse(flags_recommend_frend == "POSITIVE", 1, 0),

    outlook_num = case_when(
      flags_business_outlook == "POSITIVE" ~ 1,
      flags_business_outlook == "NEUTRAL"  ~ 0.5,
      flags_business_outlook == "NEGATIVE" ~ 0,
      TRUE ~ NA_real_
    ),

    ceo_num = case_when(
      flags_ceo_approval == "APPROVE"    ~ 1,
      flags_ceo_approval == "NO_OPINION" ~ 0.5,
      flags_ceo_approval == "DISAPPROVE" ~ 0,
      TRUE ~ NA_real_
    )
  )
```

Clustering methods need numeric inputs.
This is reasonable because:

Positive sentiment = larger value
Negative sentiment = smaller value
Neutral or no opinion = middle value

### Aggregate the data to the company level
```{r}
company_data <- glassdoor_model %>%
  group_by(company_name) %>%
  summarise(
    work_life = mean(rating_work_life, na.rm = TRUE),
    culture = mean(rating_culture_values, na.rm = TRUE),
    leadership = mean(rating_senior_leadership, na.rm = TRUE),
    career = mean(rating_career_opportunities, na.rm = TRUE),
    compensation = mean(rating_compensation_benefits, na.rm = TRUE),
    overall = mean(rating_overall, na.rm = TRUE),
    diversity = mean(rating_diversity_inclusion, na.rm = TRUE),
    recommend = mean(recommend_num, na.rm = TRUE),
    outlook = mean(outlook_num, na.rm = TRUE),
    ceo_approval = mean(ceo_num, na.rm = TRUE),
    review_count = n()
  )

nrow(company_data)
```

The data is currently at the review level, meaning each row is one employee 
review.

But the question is about similar companies, so the clustering must happen at 
the company level.

This block creates one row per company, where each feature is the average across
that company’s reviews.

### Remove companies with too few reviews
```{r}
company_data_filtered <- company_data %>%
  filter(review_count >= 2)

# Check how many companies remain
nrow(company_data_filtered)
```

A minimum review threshold was applied to ensure stability in company-level 
averages. While higher thresholds (e.g., 5 reviews) produce more reliable 
estimates, they significantly reduce the number of companies available for 
analysis. A threshold of at least 2 reviews was selected to balance reliability
and sample size, resulting in 72 companies included in the clustering analysis.
This choice allows for more meaningful cluster structure while acknowledging
some variability in companies with fewer reviews.

### Create the feature matrix for clustering
```{r}
clustering_features <- company_data_filtered %>%
  select(
    work_life,
    culture,
    leadership,
    career,
    compensation,
    overall,
    diversity,
    recommend,
    outlook,
    ceo_approval
  )
```

### Scale the Features
```{r}
scaled_features <- scale(clustering_features)

# Look at first few rows
head(scaled_features)
```

### Feature Correlation Analysis
```{r}
feature_cor <- cor(clustering_features)

pheatmap(
  feature_cor,
  main = "Correlation Heatmap of Culture Features",
  display_numbers = TRUE
)
```

The correlation heatmap shows that many workplace culture variables are
moderately to strongly positively correlated. In particular, culture, 
leadership, and overall ratings exhibit strong relationships, suggesting that 
they capture a shared underlying dimension of workplace quality.

In contrast, variables such as business outlook and CEO approval are less 
strongly correlated with the core culture metrics, indicating that they may 
capture distinct aspects of employee perception.

This pattern helps explain why clustering produces a strong primary separation,
as many features move together, while still allowing for meaningful variation
across companies.

### Elbow Method
```{r}
set.seed(123)

fviz_nbclust(scaled_features, kmeans, method = "wss") +
  labs(title = "Elbow Method for Choosing Number of Clusters")
```

### Silhouette Score
```{r}
fviz_nbclust(scaled_features, kmeans, method = "silhouette") +
  labs(title = "Silhouette Method for Choosing Number of Clusters")
```

### Compare multiple k values
```{r}
set.seed(123)

k_grid <- 2:4

k_results <- purrr::map_dfr(k_grid, function(k) {
  km <- kmeans(scaled_features, centers = k, nstart = 25)
  sil_obj <- silhouette(km$cluster, dist(scaled_features))
  
  tibble(
    k = k,
    tot_withinss = km$tot.withinss,
    betweenss_ratio = km$betweenss / km$totss,
    avg_silhouette = mean(sil_obj[, 3])
  )
})

k_results
```

A comparison of clustering solutions for k = 2, 3, and 4 shows that the 
two-cluster solution achieves the highest silhouette score, indicating the
strongest separation between groups. However, it explains less variance in the
data, as reflected by the lower between-cluster sum of squares ratio.

The three-cluster solution provides a balance between cluster separation and explanatory power. While its silhouette score is slightly lower, it captures a
larger proportion of the variance and produces a more interpretable segmentation
into low, moderate, and high workplace culture groups.

The four-cluster solution further increases explained variance but results in a noticeable drop in silhouette score, suggesting over-segmentation and less
stable clusters.

Based on this trade-off, the three-cluster solution was selected as the final
model.

# 2. K-means Model
```{r}
set.seed(123)

kmeans_model <- kmeans(scaled_features, centers = 3, nstart = 25)

company_data_filtered$cluster <- factor(kmeans_model$cluster)

# View cluster sizes
table(company_data_filtered$cluster)
```

### Cluster quality metrics
```{r}
# Silhouette scores
sil <- silhouette(kmeans_model$cluster, dist(scaled_features))

# Average silhouette width
avg_sil_width <- mean(sil[, 3])
avg_sil_width

# Between-cluster sum of squares ratio
between_total_ratio <- kmeans_model$betweenss / kmeans_model$totss
between_total_ratio
```
Overall, the clustering results indicate moderate structure in the data. While
the silhouette score suggests that cluster separation is not very strong, the between-to-total sum of squares ratio shows that a substantial portion of the
variation is captured by the clustering. This implies that the identified 
clusters provide a useful, though not perfectly distinct, grouping of companies
based on workplace culture.

The relatively low silhouette score reflects the continuous nature of workplace culture, where companies may share overlapping characteristics rather than 
forming sharply separated groups.

### Silhouette plot for the final model
```{r}
fviz_silhouette(sil) +
  labs(title = "Silhouette Plot for Final K-Means Model")
```
The silhouette plot provides a company-level view of cluster membership quality. Companies with larger silhouette values are well matched to their assigned 
cluster, while companies with values near zero lie closer to the boundary 
between clusters.

### Looking at Cluster Centers
```{r}
cluster_summary <- company_data_filtered %>%
  group_by(cluster) %>%
  summarise(
    work_life = mean(work_life),
    culture = mean(culture),
    leadership = mean(leadership),
    career = mean(career),
    compensation = mean(compensation),
    overall = mean(overall),
    diversity = mean(diversity),
    recommend = mean(recommend),
    outlook = mean(outlook),
    ceo_approval = mean(ceo_approval),
    companies_in_cluster = n()
  )

cluster_summary_round <- cluster_summary %>%
  mutate(across(where(is.numeric), ~ round(., 2)))

cluster_summary_round
```

The three clusters can be interpreted as broad workplace culture groups.

Cluster 1 represents low-rated companies. These companies have the weakest 
scores across work-life balance, culture and values, leadership, and overall satisfaction. This cluster reflects less favorable employee experiences.

Cluster 2 represents moderate companies. They have mid-range scores across most culture-related dimensions. They appear to provide average employee experiences,
with neither especially poor nor especially strong workplace culture outcomes.

Cluster 3 represents high-rated companies. These companies score highest on 
work-life balance, culture and values, leadership, compensation, and overall satisfaction. This cluster captures firms with the most favorable 
employee-reported culture profiles.

### Visualize the Clusters
```{r}
fviz_cluster(kmeans_model, data = scaled_features,
             geom = "point",
             ellipse.type = "convex",
             main = "K-Means Clustering of Companies by Culture")
```

### PCA Plot
```{r}
pca_df <- prcomp(scaled_features)

pca_scores <- as.data.frame(pca_df$x[, 1:2])
pca_scores$company_name <- company_data_filtered$company_name
pca_scores$cluster <- company_data_filtered$cluster

highlight_companies <- c("Google", "Microsoft", "Apple", "Amazon", "Meta")

pca_scores$label <- ifelse(
  pca_scores$company_name %in% highlight_companies,
  pca_scores$company_name,
  NA
)

ggplot(pca_scores, aes(x = PC1, y = PC2, color = cluster)) +
  geom_point(size = 3, alpha = 0.8) +
  geom_text_repel(aes(label = label), na.rm = TRUE) +
  labs(
    title = "PCA View of Company Culture Clusters",
    subtitle = "Selected companies labeled",
    x = "PC1",
    y = "PC2"
  )
```

# Cluster Profile Plot
```{r}
cluster_profile_long <- cluster_summary %>%
  select(cluster, work_life, culture, leadership, career,
         compensation, overall, diversity, recommend, outlook, ceo_approval) %>%
  pivot_longer(
    cols = -cluster,
    names_to = "feature",
    values_to = "mean_value"
  )

ggplot(cluster_profile_long, aes(x = feature, y = mean_value, fill = cluster)) +
  geom_col(position = "dodge") +
  theme_minimal() +
  labs(
    title = "Cluster Profiles Across Culture Features",
    x = "Feature",
    y = "Cluster Mean"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Find companies similar to a chosen company
```{r}
target_company <- "Google"

target_cluster <- company_data_filtered %>%
  filter(company_name == target_company) %>%
  pull(cluster)

similar_companies <- company_data_filtered %>%
  filter(cluster == target_cluster) %>%
  arrange(desc(work_life), desc(culture))

similar_companies
```

### Find the most similar companies by distance
```{r}
scaled_df <- as.data.frame(scaled_features)
scaled_df$company_name <- company_data_filtered$company_name

target_company <- "Google"

target_row <- scaled_df %>%
  filter(company_name == target_company) %>%
  select(-company_name)

nearest_companies <- scaled_df %>%
  rowwise() %>%
  mutate(distance = sqrt(sum((c_across(-company_name) - as.numeric(target_row))^2))) %>%
  ungroup() %>%
  left_join(
    company_data_filtered %>%
      select(company_name, cluster, overall, culture, work_life, review_count),
    by = "company_name"
  ) %>%
  arrange(distance) %>%
  slice(1:10)

nearest_companies
```

```{r}
ggplot(nearest_companies %>% filter(company_name != target_company),
       aes(x = reorder(company_name, distance), y = distance, fill = cluster)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Companies Most Similar to Google",
    x = "Company",
    y = "Distance"
  )
```

### Feature Contribution to Similarity
```{r}
# Compute per-feature squared differences
feature_contributions <- scaled_df %>%
  filter(company_name != target_company) %>%
  mutate(across(-company_name,
                ~ (. - as.numeric(target_row[[cur_column()]]))^2)) %>%
  mutate(total_distance = rowSums(across(-company_name))) %>%
  arrange(total_distance)

# Top 5 most similar companies
top_similar <- feature_contributions %>%
  slice(1:5)

top_similar

# Average contribution per feature
feature_importance <- top_similar %>%
  select(-company_name, -total_distance) %>%
  summarise(across(everything(), mean)) %>%
  pivot_longer(cols = everything(),
               names_to = "feature",
               values_to = "avg_contribution") %>%
  arrange(avg_contribution)

feature_importance
```

```{r}
ggplot(feature_importance,
       aes(x = reorder(feature, avg_contribution),
           y = avg_contribution)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Feature Contributions to Similarity with Google",
    x = "Feature",
    y = "Average Squared Difference"
  )
```

To better understand what drives similarity, the Euclidean distance was 
decomposed into feature-level contributions, allowing identification of which
workplace culture dimensions are most similar between Google and its nearest
companies.

# Limitations of K-Means
This analysis has several limitations.

First, some companies have only a small number of reviews, which may make their
average culture ratings less stable. Although a minimum threshold of two reviews
was used to preserve sample size, this still introduces some uncertainty. 

Second, all selected features were treated as equally important in the 
clustering process after scaling, even though work-life balance and culture 
values may be more central to the research question than other variables. 

Third, K-means clustering requires the number of clusters to be chosen in 
advance, so the final solution depends partly on the selected value of k. 

Finally, the results reflect only the available Glassdoor reviews in this sample
and may not fully capture the broader culture of each company.

# Conclusion
Overall, the K-means clustering analysis identified three broad groups of 
companies based on employee-reported workplace culture characteristics. The 
results suggest that companies can be grouped into low, moderate, and 
high-culture profiles using features such as work-life balance, culture and 
values, leadership, compensation, and recommendation rates. By combining cluster membership with distance-based similarity, the analysis provides a pragmatic 
way to identify companies with workplace cultures similar to a target firm.

# 3. Hierarchical Clustering
```{r}
dist_matrix <- dist(scaled_features, method = "euclidean")

hc_model <- hclust(dist_matrix, method = "ward.D2")
```

### Dendrogram
```{r}
fviz_dend(
  hc_model,
  k = 3,
  rect = TRUE,
  rect_fill = TRUE,
  show_labels = FALSE,
  lwd = 0.6,
  main = "Hierarchical Clustering Dendrogram"
)
```

The dendrogram shows a large separation at higher levels of the hierarchy, 
indicating that some groups of companies are more distinctly separated, while
finer structure exists within these broader groups.

### Cut into clusters
```{r}
hc_clusters <- cutree(hc_model, k = 3)

company_data_filtered$hc_cluster <- factor(hc_clusters)

# Compare with K-means
table(KMeans = company_data_filtered$cluster,
      Hierarchical = company_data_filtered$hc_cluster)
```
```{r}
sil_hc <- silhouette(hc_clusters, dist(scaled_features))
data.frame(
  sil_h = mean(sil_hc[,3]),
  sil_k = mean(sil[,3])
)
```

To compare clustering quality across methods, silhouette scores were computed
for both K-means and hierarchical clustering. The K-means model achieved a 
slightly higher silhouette score (0.242) compared to hierarchical clustering
(0.209), indicating that it provides better separation between clusters.

A comparison between K-means and hierarchical clustering shows that the cluster assignments are largely consistent across methods. In particular, one cluster is perfectly aligned between the two approaches, indicating a stable group of 
companies with similar culture characteristics.

However, some variation exists in the middle cluster, where companies are 
distributed across multiple hierarchical groups. This suggests that companies 
with moderate workplace culture characteristics are less distinctly separated, reflecting the continuous nature of culture-related variables.

Overall, the general agreement between the two clustering methods supports the robustness of the identified company groupings.

### Hierarchical cluster scatterplot
```{r}
fviz_cluster(
  list(data = scaled_features, cluster = hc_clusters),
  geom = "point",
  ellipse.type = "convex",
  main = "Hierarchical Clustering of Companies"
)
```

The hierarchical cluster visualization shows clear separation along the first
principal dimension, which explains the majority of the variation in the data
(61.6%). Companies in the high-culture cluster are well separated from the rest,
while moderate companies show some overlap with adjacent groups. This pattern is consistent with earlier findings that workplace culture varies along a continuum
rather than forming perfectly distinct categories.

### Heatmap of clustered companies
```{r}
pheatmap(
  scale(clustering_features),
  cluster_rows = hc_model,
  cluster_cols = TRUE,
  show_rownames = FALSE,
  main = "Clustered Heatmap of Company Culture Features"
)
```

The heatmap groups companies based on similarity rather than ordering them from
highest to lowest. One group of companies shows consistently below-average 
values across most features (blue tones), representing lower workplace culture
ratings. Another group shows values closer to the average (white tones), 
indicating moderate workplace culture. A third group exhibits higher values
across features (warmer tones), representing stronger workplace culture 
profiles.

This confirms that the clustering captures meaningful differences in company 
culture, even though the ordering of groups in the heatmap is not based on
ranking.

# Limitations of Hierarchical Clustering
Hierarchical clustering also has several limitations. First, the method is
sensitive to the choice of distance metric and linkage criteria, meaning that 
different specifications (e.g., Euclidean distance or alternative 
linkage methods) may produce different cluster structures.

Second, hierarchical clustering is based on a sequence of irreversible merges,
so early grouping decisions cannot be adjusted later in the process. This can
lead to suboptimal cluster assignments if initial merges are not ideal.

Additionally, the method can be sensitive to outliers, which may distort the
dendrogram structure and influence cluster formation. Finally, unlike K-means, hierarchical clustering does not provide a clear objective criterion for 
selecting the number of clusters, requiring subjective interpretation of the dendrogram.
