---
title: "AMS597 Project Data Cleaning"
author: "Group 1"
date: "2026-03-12"
output: 
  pdf_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries
```{r}
library(dplyr)
library(stringr)
library(lubridate)
library(tidyr)
```

# Reading the File
```{r}
library(here)

df <- read.csv(
  here("data", "processed", "glassdoor_reviews_with_sector_revised.csv")
)
colSums(is.na(df))
```


# Removing duplicate reviews
```{r}
df_clean <- df %>%
  distinct(review_id, .keep_all = TRUE)

sum(duplicated(df_clean$review_id))
```

# Clean date column
```{r}
df_clean <- df_clean %>%
  mutate(
    rating_date = gsub('"', "", rating_date),
    rating_date = as.POSIXct(rating_date, format = "%Y-%m-%dT%H:%M:%OSZ", tz = "UTC"),
    rating_date = as.Date(rating_date)
  )

# Converts the timestamp into a proper date format for time-based analysis.
```


# Fixing helpful / unhelpful counts
```{r}
df_clean <- df_clean %>%
  mutate(
    count_helpful = as.numeric(ifelse(is.na(count_helpful), 0, count_helpful)),
    count_unhelpful = as.numeric(ifelse(is.na(count_unhelpful), 0, count_unhelpful))
  ) %>%
  mutate(
    helpfulness_score = count_helpful-count_unhelpful
  )

colSums(is.na(df_clean))
```

# Employee Status And Job Title
```{r}
df_clean <- df_clean %>%
  mutate(
    employee_status = ifelse(is.na(employee_status) | employee_status == "", "Unknown", employee_status),
    employee_job_title = ifelse(is.na(employee_job_title) | employee_job_title == "", "Unknown", employee_job_title)
  )

# Filling missing values with unknown
```

# Standardizing Employee Status and Job Title
```{r}
df_clean <- df_clean %>%
  mutate(
    employee_status = str_to_title(str_trim(employee_status)),
    employee_job_title = str_to_title(str_squish(employee_job_title))
  )

# Removes extra spaces and standardizes capitals
```

# Employee location
```{r}
df_clean <- df_clean %>%
  mutate(
    employee_location = ifelse(is.na(employee_location) | employee_location == "", "Unknown", employee_location),
    employee_location = str_to_title(str_squish(employee_location))
  )

# Filling missing locations to unknown and standardizes
```



# Rename and clean COVID flag
```{r}
df_clean <- df_clean %>%
  rename(covid_related_review = flag_covid) %>%
  mutate(
    covid_related_review = ifelse(is.na(covid_related_review), 
                                  ifelse(year(rating_date)<=2023 & year(rating_date)>2019,1,0),
                                  ifelse(covid_related_review == TRUE, 1, 0))
  )

# Renaming the column and converting to binary
```


# Converting numeric columns
```{r}
df_clean <- df_clean %>%
  mutate(
    employee_job_end_year = as.numeric(employee_job_end_year),
    employee_length = as.numeric(employee_length),
  ) %>% 
  mutate(
    rating_culture_values = as.numeric(rating_culture_values),
    rating_diversity_inclusion = as.numeric(rating_diversity_inclusion),
    rating_overall = as.numeric(rating_overall),
    rating_work_life = as.numeric(rating_work_life),
    rating_compensation_benefits = as.numeric(rating_compensation_benefits),
    rating_senior_leadership = as.numeric(rating_senior_leadership),
    rating_career_opportunities = as.numeric(rating_career_opportunities)
  )

```

# Impute rating_diversity_inclusion by company median
```{r}
df_clean <- df_clean %>%
  group_by(company_name) %>%
  mutate(
    rating_diversity_inclusion = ifelse(
      is.na(rating_diversity_inclusion),
      median(rating_diversity_inclusion, na.rm = TRUE),
      rating_diversity_inclusion
    )
  ) %>%
  ungroup()
```

```{r}
sum(is.na(df_clean$rating_diversity_inclusion))
```

# Fill any remaining diversity missing values with overall median
```{r}
overall_div_median <- median(df_clean$rating_diversity_inclusion, na.rm = TRUE)

df_clean <- df_clean %>%
  mutate(
    rating_diversity_inclusion = ifelse(
      is.na(rating_diversity_inclusion),
      overall_div_median,
      rating_diversity_inclusion
    )
  )

# Uses the overall dataset median if company-level data is unavailable.
```

```{r}
sum(is.na(df_clean$rating_diversity_inclusion))
```

# median imputation for other numeric columns
```{r}
df_clean <- df_clean %>%
  mutate(
    employee_length = ifelse(
      is.na(employee_length),
      median(employee_length, na.rm = TRUE),
      employee_length
    ),
    rating_career_opportunities = ifelse(
      is.na(rating_career_opportunities),
      median(rating_career_opportunities, na.rm = TRUE),
      rating_career_opportunities
    )
  ) %>% 
  mutate(
    employee_job_quit = ifelse(grepl("^Former", employee_type, ignore.case = TRUE),TRUE, FALSE),
    employee_job_start_year =  ifelse(employee_job_quit,
        employee_job_end_year - employee_length, # Former use the year quit
        year(rating_date) - employee_length
    )
  )
```

# Clean flag columns
```{r}
df_clean <- df_clean %>%
  mutate(
    flags_business_outlook = ifelse(is.na(flags_business_outlook) | flags_business_outlook == "", "Unknown", flags_business_outlook),
    flags_ceo_approval = ifelse(is.na(flags_ceo_approval) | flags_ceo_approval == "", "Unknown", flags_ceo_approval),
    flags_recommend_frend = ifelse(is.na(flags_recommend_frend) | flags_recommend_frend == "", "Unknown", flags_recommend_frend)
  )

df_clean <- df_clean %>%
  mutate(
    flags_business_outlook = str_to_title(str_trim(flags_business_outlook)),
    flags_ceo_approval = str_to_title(str_trim(flags_ceo_approval)),
    flags_recommend_frend = str_to_title(str_trim(flags_recommend_frend))
  )

# Replaces missing values with "Unknown" and standardizes formatting.
```

# Clean text columns
```{r}
df_clean <- df_clean %>%
  mutate(
    summary = replace_na(summary, ""),
    review_pros = replace_na(review_pros, ""),
    review_cons = replace_na(review_cons, ""),
    advice_to_management = replace_na(advice_to_management, "")
  )

df_clean <- df_clean %>%
  mutate(
    summary = str_replace_all(summary, "[\r\n]+", " "),
    review_pros = str_replace_all(review_pros, "[\r\n]+", " "),
    review_cons = str_replace_all(review_cons, "[\r\n]+", " "),
    advice_to_management = str_replace_all(advice_to_management, "[\r\n]+", " ")
  )

df_clean <- df_clean %>%
  mutate(
    summary = str_squish(summary),
    review_pros = str_squish(review_pros),
    review_cons = str_squish(review_cons),
    advice_to_management = str_squish(advice_to_management)
  )

# Removes: missing values, line breaks, extra spaces
```

# Remove impossible rating values
```{r}
df_clean <- df_clean %>%
  mutate(
    rating_culture_values = ifelse(rating_culture_values < 1 | rating_culture_values > 5, NA, rating_culture_values),
    rating_diversity_inclusion = ifelse(rating_diversity_inclusion < 1 | rating_diversity_inclusion > 5, NA, rating_diversity_inclusion),
    rating_overall = ifelse(rating_overall < 1 | rating_overall > 5, NA, rating_overall),
    rating_work_life = ifelse(rating_work_life < 1 | rating_work_life > 5, NA, rating_work_life),
    rating_compensation_benefits = ifelse(rating_compensation_benefits < 1 | rating_compensation_benefits > 5, NA, rating_compensation_benefits),
    rating_senior_leadership = ifelse(rating_senior_leadership < 1 | rating_senior_leadership > 5, NA, rating_senior_leadership),
    rating_career_opportunities = ifelse(rating_career_opportunities < 1 | rating_career_opportunities > 5, NA, rating_career_opportunities)
  )

# Ensures all ratings fall within the valid range (1–5).
```

# Drop rows missing overall rating
```{r}
df_clean <- df_clean %>%
  filter(!is.na(rating_overall))

# Removes rows where the main target variable (rating_overall) is missing.
```

# Create employment status feature
```{r}
df_clean <- df_clean %>%
  mutate(
    currently_employed = ifelse(is.na(employee_job_end_year), 1, 0)
  )
```


# FillReviews

```{r}
df_clean <- df_clean %>%
  mutate(
    summary = ifelse(is.na(summary) | summary=="", summary, ""),
    advice_to_management = ifelse(is.na(advice_to_management) | advice_to_management=="", advice_to_management, ""),
    review_pros = ifelse(is.na(review_pros) | review_pros=="", review_pros, ""),
    review_cons = ifelse(is.na(review_cons) | review_cons=="", review_cons, "")
  )
```


# Dropping Unecessary Columns
```{r}
df_clean <- df_clean %>%
  select(
    -timestamp,
    -overview_id,
    -review_url,
    -employee_responses,
    -employee_type,
    -flag_featured,
    -review_advice,
    -url,
    -overview_url,
    -original_url,
    -glassdoor_employer_id
  )
```


# Final Checks
```{r}
colSums(is.na(df_clean))
str(df_clean)
summary(df_clean)
dim(df_clean)
```



# Saving cleaned file
```{r}
write.csv(df_clean, "glassdoor_reviews_cleaned.csv", row.names = FALSE)
```
